---
title: "QCalculator2_19Feb26_JH"
author: "Jessica Herrmann"
date: "2026-02-19"
output: html_document
---

SUMMARY:

This script calculates estimated streamflow by site, for each of fourteen select sites at which pressure transducers were placed by SSI in the summer of 2025 (Delta Project). The following script uses the site-specific formulas for estimated channel wetted perimeter and cross-sectional area, derived from regression models generated in the "Channel Geom Calculator 2.0.xlsx" document available in the parent Drive folder. From there, Manning's Equation is used to estimate discharge over time. 

Streamgages used in this project were deployed from the summer of 2025 through the summer of 2026. Please note that stage data offloaded from gages must first be compensated for relative barometric pressure in HOBOware.

Also included in this script is an annotated example calculation that uses facsimile data for one of SSI's survey sites, Cat Creek. Please do not use example data for final calculations of streamflow at this site.



EXPLANATION OF VARIABLES:

S = the average channel slope

R_0 = estimated hydraulic radius at time t = 0, when the transducer was placed

n = Manning's channel roughness coefficient

h_0 = the starting depth of water over the logger

h = the recorded stage heights over time at time t

R = the hydraulic radius at time t

WP = wetted perimeter of the channel cross-section at time t

CSA = cross-sectional area of water in the channel at time t


** Units of length are in cm unless otherwise stated.



```{r Package libraries}
library(curl)
library(tidyverse)
library(tidyr)
library(stringr)
library(ggplot2)
```

```{r Load data}
## Import file with channel geom. values
Channel.geom <- curl("https://raw.githubusercontent.com/SierraStreamsInstitute/Delta/refs/heads/main/Estimated%20Streamflow/DeltaChannelGeom.csv")
Channel.geom <- read.csv(Channel.geom, header = T, sep = ",")

## Stage data for your site (may want to split this up into separate scripts by site so it doesn't get too unruly). This dataset contains baro. pressure-compensated stage height data from each pressure transducer. Barometric compensation is done in HOBOware.



#### EXAMPLE ####
# # Using dummy dataset here for Cat Crk. below:
# CAT.stage <- curl("https://raw.githubusercontent.com/jnherrmann/Streamflow_Estimation_Practice/refs/heads/main/CatCrk_FacsimileTestData.csv")
# CAT.stage <- read.csv(CAT.stage, header = T, sep = ",")
# # CAT.stage <- CAT.stage[,c(1:6,11:12)]
# # This example dataset is saved in JH's personal GitHub (publicly available), so as not to confuse with real datasets.



#### REAL STAGE DATASETS HERE ####
## You'll want to upload your compensated stage data below.I like to store datasets in GitHub and then call them using the curl() function. But up to you if you'd prefer a different method, e.g. saving locally.

SHADY.stage <- curl("")
SHADY.stage <- read.csv(SHADY.stage, header = T, sep = ",")

DC116.stage <- curl("")
DC116.stage <- read.csv(DC116.stage, header = T, sep = ",")

DC20.stage <- curl("")
DC20.stage <- read.csv(DC20.stage, header = T, sep = ",")

DC10.stage <- curl("")
DC10.stage <- read.csv(DC10.stage, header = T, sep = ",")

BR55.stage <- curl("")
BR55.stage <- read.csv(BR55.stage, header = T, sep = ",")

EFAR.stage <- curl("")
EFAR.stage <- read.csv(EFAR.stage, header = T, sep = ",")

MON.stage <- curl("")
MON.stage <- read.csv(MON.stage, header = T, sep = ",")

GSW.stage <- curl("")
GSW.stage <- read.csv(GSW.stage, header = T, sep = ",")

NFFR.stage <- curl("")
NFFR.stage <- read.csv(NFFR.stage, header = T, sep = ",")

MAR.stage <- curl("")
MAR.stage <- read.csv(MAR.stage, header = T, sep = ",")

CAT.stage <- curl("")
CAT.stage <- read.csv(CAT.stage, header = T, sep = ",")

BUT.stage <- curl("")
BUT.stage <- read.csv(BUT.stage, header = T, sep = ",")

WBUT.stage <- curl("")
WBUT.stage <- read.csv(WBUT.stage, header = T, sep = ",")

MCL.stage <- curl("")
MCL.stage <- read.csv(MCL.stage, header = T, sep = ",")
```


```{r Stage Data Cleanup}
#### EXAMPLE ####
## Tidying stage dataset
# CAT.stage$Sensor_Depth_feet <- 30.48*CAT.stage$Sensor_Depth_feet # Convert "Sensor_Depth_feet" from feet to cm
# colnames(CAT.stage)[colnames(CAT.stage) == "Sensor_Depth_feet"] <- "h" # Rename as "h"
# colnames(CAT.stage)[colnames(CAT.stage) == "Date_Time"] <- "t" # Rename "Date_Time" as "t", for Time
# CAT.stage$t <- mdy_hms(CAT.stage$t) # Format as a POSIXct object rather than a character string
# CAT.geom <- Channel.geom[Channel.geom$Site == "CAT", ] # Creating a new subset for just the channel geometry values
# CAT.stage <- cbind(CAT.geom, CAT.stage) # Binding the stage data and channel geometry datasets



### REAL DATA HERE ####

## SHADY
SHADY.stage$Sensor_Depth_feet <- 30.48*SHADY.stage$Sensor_Depth_feet
colnames(SHADY.stage)[colnames(SHADY.stage) == "Sensor_Depth_feet"] <- "h"
colnames(SHADY.stage)[colnames(SHADY.stage) == "Date_Time"] <- "t"
SHADY.stage$t <- mdy_hms(SHADY.stage$t)
SHADY.geom <- Channel.geom[Channel.geom$Site == "SHADY", ]
SHADY.stage <- cbind(SHADY.geom, SHADY.stage)

## DC116
DC116.stage$Sensor_Depth_feet <- 30.48*DC116.stage$Sensor_Depth_feet
colnames(DC116.stage)[colnames(DC116.stage) == "Sensor_Depth_feet"] <- "h"
colnames(DC116.stage)[colnames(DC116.stage) == "Date_Time"] <- "t"
DC116.stage$t <- mdy_hms(DC116.stage$t)
DC116.geom <- Channel.geom[Channel.geom$Site == "DC116", ]
DC116.stage <- cbind(DC116.geom, DC116.stage)

## DC20
DC20.stage$Sensor_Depth_feet <- 30.48*DC20.stage$Sensor_Depth_feet
colnames(DC20.stage)[colnames(DC20.stage) == "Sensor_Depth_feet"] <- "h"
colnames(DC20.stage)[colnames(DC20.stage) == "Date_Time"] <- "t"
DC20.stage$t <- mdy_hms(DC20.stage$t)
DC20.geom <- Channel.geom[Channel.geom$Site == "DC20", ]
DC20.stage <- cbind(DC20.geom, DC20.stage)

## DC10
DC10.stage$Sensor_Depth_feet <- 30.48*DC10.stage$Sensor_Depth_feet
colnames(DC10.stage)[colnames(DC10.stage) == "Sensor_Depth_feet"] <- "h"
colnames(DC10.stage)[colnames(DC10.stage) == "Date_Time"] <- "t"
DC10.stage$t <- mdy_hms(DC10.stage$t)
DC10.geom <- Channel.geom[Channel.geom$Site == "DC10", ]
DC10.stage <- cbind(DC10.geom, DC10.stage)

## BR55
BR55.stage$Sensor_Depth_feet <- 30.48*BR55.stage$Sensor_Depth_feet
colnames(BR55.stage)[colnames(BR55.stage) == "Sensor_Depth_feet"] <- "h"
colnames(BR55.stage)[colnames(BR55.stage) == "Date_Time"] <- "t"
BR55.stage$t <- mdy_hms(BR55.stage$t)
BR55.geom <- Channel.geom[Channel.geom$Site == "BR55", ]
BR55.stage <- cbind(BR55.geom, BR55.stage)

## EFAR
EFAR.stage$Sensor_Depth_feet <- 30.48*EFAR.stage$Sensor_Depth_feet
colnames(EFAR.stage)[colnames(EFAR.stage) == "Sensor_Depth_feet"] <- "h"
colnames(EFAR.stage)[colnames(EFAR.stage) == "Date_Time"] <- "t"
EFAR.stage$t <- mdy_hms(EFAR.stage$t)
EFAR.geom <- Channel.geom[Channel.geom$Site == "EFAR", ]
EFAR.stage <- cbind(EFAR.geom, EFAR.stage)

## MON
MON.stage$Sensor_Depth_feet <- 30.48*MON.stage$Sensor_Depth_feet
colnames(MON.stage)[colnames(MON.stage) == "Sensor_Depth_feet"] <- "h"
colnames(MON.stage)[colnames(MON.stage) == "Date_Time"] <- "t"
MON.stage$t <- mdy_hms(MON.stage$t)
MON.geom <- Channel.geom[Channel.geom$Site == "MON", ]
MON.stage <- cbind(MON.geom, MON.stage)

## GSW
GSW.stage$Sensor_Depth_feet <- 30.48*GSW.stage$Sensor_Depth_feet
colnames(GSW.stage)[colnames(GSW.stage) == "Sensor_Depth_feet"] <- "h"
colnames(GSW.stage)[colnames(GSW.stage) == "Date_Time"] <- "t"
GSW.stage$t <- mdy_hms(GSW.stage$t)
GSW.geom <- Channel.geom[Channel.geom$Site == "GSW", ]
GSW.stage <- cbind(GSW.geom, GSW.stage)

## NFFR
NFFR.stage$Sensor_Depth_feet <- 30.48*NFFR.stage$Sensor_Depth_feet
colnames(NFFR.stage)[colnames(NFFR.stage) == "Sensor_Depth_feet"] <- "h"
colnames(NFFR.stage)[colnames(NFFR.stage) == "Date_Time"] <- "t"
NFFR.stage$t <- mdy_hms(NFFR.stage$t)
NFFR.geom <- Channel.geom[Channel.geom$Site == "NFFR", ]
NFFR.stage <- cbind(NFFR.geom, NFFR.stage)

## MAR
MAR.stage$Sensor_Depth_feet <- 30.48*MAR.stage$Sensor_Depth_feet
colnames(MAR.stage)[colnames(MAR.stage) == "Sensor_Depth_feet"] <- "h"
colnames(MAR.stage)[colnames(MAR.stage) == "Date_Time"] <- "t"
MAR.stage$t <- mdy_hms(MAR.stage$t)
MAR.geom <- Channel.geom[Channel.geom$Site == "MAR", ]
MAR.stage <- cbind(MAR.geom, MAR.stage)

## CAT
MAR.stage$Sensor_Depth_feet <- 30.48*MAR.stage$Sensor_Depth_feet
colnames(MAR.stage)[colnames(MAR.stage) == "Sensor_Depth_feet"] <- "h"
colnames(MAR.stage)[colnames(MAR.stage) == "Date_Time"] <- "t"
MAR.stage$t <- mdy_hms(MAR.stage$t)
MAR.geom <- Channel.geom[Channel.geom$Site == "MAR", ]
MAR.stage <- cbind(MAR.geom, MAR.stage)

## BUT
BUT.stage$Sensor_Depth_feet <- 30.48*BUT.stage$Sensor_Depth_feet
colnames(BUT.stage)[colnames(BUT.stage) == "Sensor_Depth_feet"] <- "h"
colnames(BUT.stage)[colnames(BUT.stage) == "Date_Time"] <- "t"
BUT.stage$t <- mdy_hms(BUT.stage$t)
BUT.geom <- Channel.geom[Channel.geom$Site == "BUT", ]
BUT.stage <- cbind(BUT.geom, BUT.stage)

## WBUT
WBUT.stage$Sensor_Depth_feet <- 30.48*WBUT.stage$Sensor_Depth_feet
colnames(WBUT.stage)[colnames(WBUT.stage) == "Sensor_Depth_feet"] <- "h"
colnames(WBUT.stage)[colnames(WBUT.stage) == "Date_Time"] <- "t"
WBUT.stage$t <- mdy_hms(WBUT.stage$t)
WBUT.geom <- Channel.geom[Channel.geom$Site == "WBUT", ]
WBUT.stage <- cbind(WBUT.geom, WBUT.stage)

## MCL
MCL.stage$Sensor_Depth_feet <- 30.48*MCL.stage$Sensor_Depth_feet
colnames(MCL.stage)[colnames(MCL.stage) == "Sensor_Depth_feet"] <- "h"
colnames(MCL.stage)[colnames(MCL.stage) == "Date_Time"] <- "t"
MCL.stage$t <- mdy_hms(MCL.stage$t)
MCL.geom <- Channel.geom[Channel.geom$Site == "MCL", ]
MCL.stage <- cbind(MCL.geom, MCL.stage)
```

```{r Calculating the hydraulic radius at time t = 0, R(0)}
## I used R to calculate (estimated) hydraulic radius for each site at t = 0.
## For the purposes of solving for Q, you can ignore this chunk because these R(0) calculations
## were used to solve for n, which is already in the ChanGeom dataset.

## SHADY 
# h(0) = 19
# Solve for hydraulic radius, R = CSA / WP:
# (1.3076*19^2 + 73.342*19 + 447.54) / (-0.0436*19^2 + 12.034*19 + 35.957)
# 
# # R(0) = 9.294583
# 
# 
# ## DC116
# # h(0) = 32
# (0.0002*32^4 - 0.0139*32^3 + 0.4598*32^2 + 513.49*32 + 23665) / (-0.0236*32^2 + 4.6381*32 + 534.69)
# 
# # R(0) = 61.19159
# 
# 
# 
# ## DC20
# # h(0) = 58
# (-0.0885*58^3 + 14.053*58^2 + 281.56*58 + 842.56) / (12.211*58 + 338.14)
# 
# # R(0) = 45.08879
# 
# 
# 
# ## DC10
# # h(0) = 23
# (0.0079*23^3 + 2.4656*23^2 + 247.86*23 + 8289) / (0.00007*23^3 + 0.0203*23^2 + 5.0659*23 + 524.35)
# 
# # R(0) = 23.5881
# 
# 
# 
# ## BR55
# # h(0) = 63
# (-0.0472*63^3 + 8.4298*63^2 + 196.07*63 + 1090.7) / (-0.052*63^2 + 11.559*63 + 189.75)
# 
# # R(0) = 49.32519
# 
# 
# 
# ## EFAR
# # h(0) = 55.5
# (-0.0245*55.5^3 + 6.8857*55.5^2 + 124.37*55.5 + 673.57) / (-0.0506*55.5^2 + 12.198*55.5 + 160.69)
# 
# # R(0) = 36.0762
# 
# 
# 
# ## MON
# # h(0) = 11
# (0.006*11^4 - 0.4309*11^3 + 13.738*11^2 + 187.57*1) / (-0.0014*11^4 + 0.111*11^3 - 2.8422*11^2 + 34.473*11 + 194.9)
# 
# # R(0) = 3.816541
# 
# 
# 
# ## GSW
# # h(0) = 34
# (-0.0694*34^3 + 8.5316*34^2 - 0.8675*34 - 17.619) / (-0.1495*34^2 + 15.135*34 + 20.284)
# 
# # R(0) = 19.57652
# 
# 
# 
# ## NFFR
# # h(0) = 20
# (0.1035*20^3 - 5.5613*20^2 + 646.81*20 + 11607) / (-0.0003*20^3 + 0.1416*20^2 + 15.192*20 + 286.68)
# 
# # R(0) = 35.89968
# 
# 
# 
# ## MAR
# # h(0) = 35
# (0.0006*35^4 - 0.0737*35^3 + 4.6675*35^2 + 96.475*35 + 250.65) / (0.000006*35^4 - 0.0012*35^3 + 0.0791*35^2 + 6.0092*35 + 62.038)
# 
# # R(0) = 21.68056
# 
# 
# 
# ## CAT
# # h(0) = 48
# (2.237*48^2 + 285.09*48 + 7401) / (0.0013*48^3 - 0.1165*48^2 + 4.029*48 + 401.13)
# 
# # R(0) = 55.84322
# 
# 
# ## BUT
# # h(0) = 34
# (5.0538*34^2 + 648.17*34 + 6798.8) / (0.0032*34^3 - 0.6757*34^2 + 42.024*34 + 652.75)
# 
# # R(0) = 24.315
# 
# 
# 
# ## WBUT
# # h(0) = 38
# (-0.0003*38^4 + 0.0445*38^3 + 1.1027*38^2 + 46.012*38 + 841.3) / (-0.00006*38^3 + 0.0126*38^2 + 3.6795*38 + 99.273)
# 
# # R(0) = 23.61579
# 
# 
# 
# ## MCL
# # h(0) = 32
# (-0.0582*32^3 + 12.539*32^2 + 323.06*32 + 2067.7) / (6.863*32 + 413.17)
# 
# # R(0) = 36.88207
```

Below are all of the regression models for WP and CSA, given stage height, h, by individual site.

SHADY:
WP.SHADY = -0.0436*h^2 + 12.034*h + 35.957
CSA.SHADY = 1.3076*h^2 + 73.342*h + 447.54

DC116:
WP.DC116 = -0.0236*h^2 + 4.6381*h + 534.69
CSA.DC116 = 0.0002*h^4 - 0.0139*h^3 + 0.4598*h^2 + 513.49*h + 23665

DC20:
WP.DC20 = 12.211*h + 338.14
CSA.DC20 = -0.0885*h^3 + 14.053*h^2 + 281.56*h + 842.56

DC10:
WP.DC10 = 0.00007*h^3 + 0.0203*h^2 + 5.0659*h + 524.35
CSA.DC10 = 0.0079*h^3 + 2.4656*h^2 + 247.86*h + 8289

BR55:
WP.BR55 = -0.052*h^2 + 11.559*h + 189.75
CSA.BR55 = -0.0472*h^3 + 8.4298*h^2 + 196.07*h + 1090.7

EFAR:
WP.EFAR = -0.0506*h^2 + 12.198*h + 160.69
CSA.EFAR = -0.0245*h^3 + 6.8857*h^2 + 124.37*h + 673.57

MON:
WP.MON = -0.0014*h^4 + 0.111*h^3 - 2.8422*h^2 + 34.473*h + 194.9
CSA.MON = 0.006*h^4 - 0.4309*h^3 + 13.738*h^2 + 187.57*h

GSW:
WP.GSW = -0.1495*h^2 + 15.135*h + 20.284
CSA.GSW = -0.0694*h^3 + 8.5316*h^2 - 0.8675*h - 17.619

NFFR:
WP.NFFR = -0.0003*h^3 + 0.1416*h^2 + 15.192*h + 286.68
CSA.NFFR = 0.1035*h^3 - 5.5613*h^2 + 646.81*h + 11607

MAR:
WP.MAR = 0.000006*h^4 - 0.0012*h^3 + 0.0791*h^2 + 6.0092*h + 62.038
CSA.MAR = 0.0006*h^4 - 0.0737*h^3 + 4.6675*h^2 + 96.475*h + 250.65

CAT:
WP.CAT = 0.0013*h^3 - 0.1165*h^2 + 4.029*h + 401.13
CSA.CAT = 2.237*h^2 + 285.09*h + 7401

BUT:
WP.BUT = 0.0032*h^3 - 0.6757*h^2 + 42.024*h + 652.75
CSA.BUT = 5.0538*h^2 + 648.17*h + 6798.8

WBUT:
WP.WBUT = -0.00006*h^3 + 0.0126*h^2 + 3.6795*h + 99.273
CSA.WBUT = -0.0003*h^4 + 0.0445*h^3 + 1.1027*h^2 + 46.012*h + 841.3

MCL:
WP.MCL = 6.863*h + 413.17
CSA.MCL = -0.0582*h^3 + 12.539*h^2 + 323.06*h + 2067.7

```{r Calculate WP, CSA, and Q}
## Next calculate WP and CSA over time, using the equation for your site of interest.

#### EXAMPLE ####
# CAT.Q <- CAT.stage %>%
#   mutate(
#     WP = 0.0013*h^3 - 0.1165*h^2 + 4.029*h + 401.13,
#     CSA = 2.237*h^2 + 285.09*h + 7401,
#     R = CSA / WP,
#     Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))



#### REAL DATA HERE ####

## SHADY
SHADY.Q <- SHADY.stage %>%
  mutate(
    WP = -0.0436*h^2 + 12.034*h + 35.957,
    CSA = 1.3076*h^2 + 73.342*h + 447.54,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## DC116
DC116.Q <- DC116.stage %>%
  mutate(
    WP = -0.0236*h^2 + 4.6381*h + 534.69,
    CSA = 0.0002*h^4 - 0.0139*h^3 + 0.4598*h^2 + 513.49*h + 23665,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## DC20
DC20.Q <- DC20.stage %>%
  mutate(
    WP = 12.211*h + 338.14,
    CSA = -0.0885*h^3 + 14.053*h^2 + 281.56*h + 842.56,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## DC10
DC10.Q <- DC10.stage %>%
  mutate(
    WP = 0.00007*h^3 + 0.0203*h^2 + 5.0659*h + 524.35,
    CSA = 0.0079*h^3 + 2.4656*h^2 + 247.86*h + 8289,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## BR55
BR55.Q <- BR55.stage %>%
  mutate(
    WP = -0.052*h^2 + 11.559*h + 189.75,
    CSA = -0.0472*h^3 + 8.4298*h^2 + 196.07*h + 1090.7,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## EFAR
EFAR.Q <- EFAR.stage %>%
  mutate(
    WP = -0.0506*h^2 + 12.198*h + 160.69,
    CSA = -0.0245*h^3 + 6.8857*h^2 + 124.37*h + 673.57,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## MON
MON.Q <- MON.stage %>%
  mutate(
    WP = -0.0014*h^4 + 0.111*h^3 - 2.8422*h^2 + 34.473*h + 194.9,
    CSA = 0.006*h^4 - 0.4309*h^3 + 13.738*h^2 + 187.57*h,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## GSW
GSW.Q <- GSW.stage %>%
  mutate(
    WP= -0.1495*h^2 + 15.135*h + 20.284,
    CSA = -0.0694*h^3 + 8.5316*h^2 - 0.8675*h - 17.619,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## NFFR
NFFR.Q <- NFFR.stage %>%
  mutate(
    WP = -0.0003*h^3 + 0.1416*h^2 + 15.192*h + 286.68,
    CSA = 0.1035*h^3 - 5.5613*h^2 + 646.81*h + 11607,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## MAR
MAR.Q <- MAR.stage %>%
  mutate(
    WP = 0.000006*h^4 - 0.0012*h^3 + 0.0791*h^2 + 6.0092*h + 62.038,
    CSA = 0.0006*h^4 - 0.0737*h^3 + 4.6675*h^2 + 96.475*h + 250.65,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## CAT
CAT.Q <- CAT.stage %>%
  mutate(
    WP = 0.0013*h^3 - 0.1165*h^2 + 4.029*h + 401.13,
    CSA = 2.237*h^2 + 285.09*h + 7401,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## BUT
BUT.Q <- BUT.stage %>%
  mutate(
    WP = 0.0032*h^3 - 0.6757*h^2 + 42.024*h + 652.75,
    CSA = 5.0538*h^2 + 648.17*h + 6798.8,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## WBUT
WBUT.Q <- WBUT.stage %>%
  mutate(
    WP = -0.00006*h^3 + 0.0126*h^2 + 3.6795*h + 99.273,
    CSA = -0.0003*h^4 + 0.0445*h^3 + 1.1027*h^2 + 46.012*h + 841.3,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))

## MCL
MCL.Q <- MCL.stage %>%
  mutate(
    WP = 6.863*h + 413.17,
    CSA = -0.0582*h^3 + 12.539*h^2 + 323.06*h + 2067.7,
    R = CSA / WP,
    Q = (1.49/n)*(CSA)*(R^(2/3))*(S^(1/2)))
```

```{r Final Cleanup}
## NOTE on UNITS - WP and CSA were calculated in cm and cm^2, respectively. Therefore, R is also in cm.
## n is in sec^-1 (weirdly, but it works). Average channel slope, S, is a unitless decimal value.

#### EXAMPLE ####
## Simplified df with just time (t) and flow (Q)
# CAT.Q2 <- subset(CAT.Q, select = c("t", "Q")) # Select columns
# colnames(CAT.Q2)[colnames(CAT.Q2) == "Q"] <- "Q_ccs" # Specifying units; "CCS" = cubic cm/sec
# 
# ## Convert Q_t from cubic cm/sec to cubic feet/sec (cfs)
# CAT.Q2 <- CAT.Q2 %>%
#   mutate(Q_cfs = Q_ccs*0.0000353146848166)



#### REAL DATA HERE ####

## SHADY
SHADY.Q2 <- subset(SHADY.Q, select = c("t", "Q"))
colnames(SHADY.Q2)[colnames(SHADY.Q2) == "Q"] <- "Q_ccs"
SHADY.Q2 <- SHADY.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## DC116
DC116.Q2 <- subset(DC116.Q, select = c("t", "Q"))
colnames(DC116.Q2)[colnames(DC116.Q2) == "Q"] <- "Q_ccs"
DC116.Q2 <- DC116.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## DC20
DC20.Q2 <- subset(DC20.Q, select = c("t", "Q"))
colnames(DC20.Q2)[colnames(DC20.Q2) == "Q"] <- "Q_ccs"
DC20.Q2 <- DC20.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## DC10
DC10.Q2 <- subset(DC10.Q, select = c("t", "Q"))
colnames(DC10.Q2)[colnames(DC10.Q2) == "Q"] <- "Q_ccs"
DC10.Q2 <- DC10.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## BR55
BR55.Q2 <- subset(BR55.Q, select = c("t", "Q"))
colnames(BR55.Q2)[colnames(BR55.Q2) == "Q"] <- "Q_ccs"
BR55.Q2 <- BR55.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## EFAR
EFAR.Q2 <- subset(EFAR.Q, select = c("t", "Q"))
colnames(EFAR.Q2)[colnames(EFAR.Q2) == "Q"] <- "Q_ccs"
EFAR.Q2 <- EFAR.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## MON
MON.Q2 <- subset(MON.Q, select = c("t", "Q"))
colnames(MON.Q2)[colnames(MON.Q2) == "Q"] <- "Q_ccs"
MON.Q2 <- MON.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## GSW
GSW.Q2 <- subset(GSW.Q, select = c("t", "Q"))
colnames(GSW.Q2)[colnames(GSW.Q2) == "Q"] <- "Q_ccs"
GSW.Q2 <- GSW.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## NFFR
NFFR.Q2 <- subset(NFFR.Q, select = c("t", "Q"))
colnames(NFFR.Q2)[colnames(NFFR.Q2) == "Q"] <- "Q_ccs"
NFFR.Q2 <- NFFR.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## MAR
MAR.Q2 <- subset(MAR.Q, select = c("t", "Q"))
colnames(MAR.Q2)[colnames(MAR.Q2) == "Q"] <- "Q_ccs"
MAR.Q2 <- MAR.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## CAT
CAT.Q2 <- subset(CAT.Q, select = c("t", "Q"))
colnames(CAT.Q2)[colnames(CAT.Q2) == "Q"] <- "Q_ccs"
CAT.Q2 <- CAT.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## BUT
BUT.Q2 <- subset(BUT.Q, select = c("t", "Q"))
colnames(BUT.Q2)[colnames(CAT.Q2) == "Q"] <- "Q_ccs"
BUT.Q2 <- BUT.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## WBUT
WBUT.Q2 <- subset(WBUT.Q, select = c("t", "Q"))
colnames(WBUT.Q2)[colnames(WCAT.Q2) == "Q"] <- "Q_ccs"
WBUT.Q2 <- WBUT.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)

## MCL
MCL.Q2 <- subset(MCL.Q, select = c("t", "Q"))
colnames(MCL.Q2)[colnames(CAT.Q2) == "Q"] <- "Q_ccs"
MCL.Q2 <- MCL.Q2 %>%
  mutate(Q_cfs = Q_ccs*0.0000353146848166)
```

```{r Testing the Results}
# ## Does this produce a believable hydrograph?
# 
# ## Estimated hydrograph
# Qest.plot <- ggplot(CAT.Q2, aes(x = t, y = Q_cfs)) +
#                   geom_line()
# Qest.plot
# 
# ## Stage data for comparison
# stage.plot <- ggplot(CAT.Q, aes(x = t, y = h)) +
#                   geom_line()
# stage.plot
# 
# ## For fun, you can also look at the plots for WP and CSA:
# WP.plot1 <- ggplot(CAT.Q, aes(x = t, y = WP)) +
#                   geom_line()
# WP.plot1
# 
# CSA.plot1 <- ggplot(CAT.Q, aes(x = t, y = CSA)) +
#                   geom_line()
# CSA.plot1
# 
# ## Lastly, here are the original regression models (built in Excel) for approximating WP and CSA from stage height:
# WP.plot2 <- ggplot(CAT.Q, aes(x = h, y = WP)) +
#                   geom_line()
# WP.plot2
# 
# CSA.plot2 <- ggplot(CAT.Q, aes(x = h, y = CSA)) +
#                   geom_line()
# CSA.plot2
```

```{r Export to a CSV}
#### EXAMPLE ####
## Export complete estimated flow dataset to a CSV file
# write.csv(CAT.Q2, "C:\\Users\\jessicaherrmann\\Desktop\\Delta_CatCrk_EstimatedFlows.csv")
```
