---
title: "Delta_NLCDEdit_JHerrmann_2026"
author: "Jessica Herrmann"
date: "2026-05-06"
output: html_document
---

This script reformats the output tables generated in the DeltaNLCDData ArcGIS project:
- Combining 1-km and 5-km buffered datasets
- Converting land cover class codes to brief descriptions
- The buffer radius around each site point is also included as a modifier for land cover column headers. I organized the data this way so that it would be easier to use for NMDS later on.

```{r Packages}
library(dplyr)
library(tidyr)
library(tidyverse)
library(lubridate)
```

```{r Datasets}
## NLCD Land Cover data from 2004 - 2024, within a 1-km buffer around each site
LC.1km <- LndCov20042024_1KM

## NLCD Land Cover data from 2004 - 2024, within a 5-km buffer around each site
LC.5km <- LndCov20042024_5KM
```

```{r Rename LC Classes}
#### 1-km buffered dataset ####
## Reformatting
LC.1km <- LC.1km[,c(1,6,3:4)]
names(LC.1km)[names(LC.1km) == "Value"] <- "LndCov.1km"
names(LC.1km)[names(LC.1km) == "Count"] <- "PixelCount.1km"

## Renaming LC Classes
## Reference: <https://mnatlas.org/metadata/nlcd_2021_landcover_UTM.html>
LC.1km <- LC.1km %>%
  mutate(LndCov.1km = recode(LndCov.1km,
                        "0" = "Unclassified.1km", 
                        "11" = "OpenWater.1km",
                        "12" = "PerennialIceSnow.1km",
                        "21" = "DevOpenSpace.1km",
                        "22" = "DevLowIntensity.1km",
                        "23" = "DevMedIntensity.1km",
                        "24" = "DevHighIntensity.1km",
                        "31" = "BarrenLand.1km",
                        "41" = "DeciduousForest.1km",
                        "42" = "EvergreenForest.1km",
                        "43" = "MixedForest.1km",
                        "51" = "AKDwarfScrub.1km",
                        "52" = "Scrub.1km",
                        "71" = "GrasslandHerb.1km",
                        "72" = "SedgeHerb.1km",
                        "73" = "AKLichens.1km",
                        "74" = "AKMoss.1km",
                        "81" = "PastureHay.1km",
                        "82" = "CultivCrops.1km",
                        "90" = "WoodyWetlands.1km",
                        "95" = "EmergHerbWetlands.1km"
                        ))

## Pivot wider on LndCov column
LC.1km.wide <- LC.1km %>%
  group_by(Site, Year) %>%
  pivot_wider(names_from = LndCov.1km, values_from = PixelCount.1km)

## Replacing NA's with 0's
LC.1km.wide[is.na(LC.1km.wide)] <- 0

## Add column with total pixel count per site/year combination
LC.1km.wide$TotalPixels.1km <- rowSums(LC.1km.wide[, 3:17], na.rm = TRUE)

## Re-calculate LC pixel counts as percentages out of the total pixel count
LC.1km.final <- LC.1km.wide %>%
  group_by(Site, Year) %>%
  mutate(
    DevOpenSpace.1km.pct = DevOpenSpace.1km / TotalPixels.1km,
    DevLowIntensity.1km.pct = DevLowIntensity.1km / TotalPixels.1km,
    DeciduousForest.1km.pct = DeciduousForest.1km / TotalPixels.1km,
    EvergreenForest.1km.pct = EvergreenForest.1km / TotalPixels.1km,
    MixedForest.1km.pct = MixedForest.1km / TotalPixels.1km,
    Scrub.1km.pct = Scrub.1km / TotalPixels.1km,
    GrasslandHerb.1km.pct = GrasslandHerb.1km / TotalPixels.1km,
    WoodyWetlands.1km.pct = WoodyWetlands.1km / TotalPixels.1km,
    EmergHerbWetlands.1km.pct = EmergHerbWetlands.1km / TotalPixels.1km,
    DevMedIntensity.1km.pct = DevMedIntensity.1km / TotalPixels.1km,
    BarrenLand.1km.pct = BarrenLand.1km / TotalPixels.1km,
    OpenWater.1km.pct = OpenWater.1km / TotalPixels.1km,
    CultivCrops.1km.pct = CultivCrops.1km / TotalPixels.1km,
    DevHighIntensity.1km.pct = DevHighIntensity.1km / TotalPixels.1km,
    PastureHay.1km.pct = PastureHay.1km / TotalPixels.1km
  )



#### 5-km buffered dataset ####
LC.5km <- LC.5km[,c(1,6,3:4)]
names(LC.5km)[names(LC.5km) == "Value"] <- "LndCov.5km"
names(LC.5km)[names(LC.5km) == "Count"] <- "PixelCount.5km"

LC.5km <- LC.5km %>%
  mutate(LndCov.5km = recode(LndCov.5km,
                        "0" = "Unclassified.5km", 
                        "11" = "OpenWater.5km",
                        "12" = "PerennialIceSnow.5km",
                        "21" = "DevOpenSpace.5km",
                        "22" = "DevLowIntensity.5km",
                        "23" = "DevMedIntensity.5km",
                        "24" = "DevHighIntensity.5km",
                        "31" = "BarrenLand.5km",
                        "41" = "DeciduousForest.5km",
                        "42" = "EvergreenForest.5km",
                        "43" = "MixedForest.5km",
                        "51" = "AKDwarfScrub.5km",
                        "52" = "Scrub.5km",
                        "71" = "GrasslandHerb.5km",
                        "72" = "SedgeHerb.5km",
                        "73" = "AKLichens.5km",
                        "74" = "AKMoss.5km",
                        "81" = "PastureHay.5km",
                        "82" = "CultivCrops.5km",
                        "90" = "WoodyWetlands.5km",
                        "95" = "EmergHerbWetlands.5km"
                        ))

LC.5km.wide <- LC.5km %>%
  group_by(Site, Year) %>%
  pivot_wider(names_from = LndCov.5km, values_from = PixelCount.5km)

LC.5km.wide[is.na(LC.5km.wide)] <- 0

LC.5km.wide$TotalPixels.5km <- rowSums(LC.5km.wide[, 3:18], na.rm = TRUE)

# names(LC.5km.wide)

LC.5km.final <- LC.5km.wide %>%
  group_by(Site, Year) %>%
  mutate(
    OpenWater.5km.pct = OpenWater.5km / TotalPixels.5km,
    DevOpenSpace.5km.pct = DevOpenSpace.5km / TotalPixels.5km,
    DevLowIntensity.5km.pct = DevLowIntensity.5km / TotalPixels.5km,
    DevMedIntensity.5km.pct = DevMedIntensity.5km / TotalPixels.5km,
    DevHighIntensity.5km.pct = DevHighIntensity.5km / TotalPixels.5km,
    BarrenLand.5km.pct = BarrenLand.5km / TotalPixels.5km,
    DeciduousForest.5km.pct = DeciduousForest.5km / TotalPixels.5km,
    EvergreenForest.5km.pct = EvergreenForest.5km / TotalPixels.5km,
    MixedForest.5km.pct = MixedForest.5km / TotalPixels.5km,
    Scrub.5km.pct = Scrub.5km / TotalPixels.5km,
    GrasslandHerb.5km.pct = GrasslandHerb.5km / TotalPixels.5km,
    PastureHay.5km.pct = PastureHay.5km / TotalPixels.5km,
    WoodyWetlands.5km.pct = WoodyWetlands.5km / TotalPixels.5km,
    EmergHerbWetlands.5km.pct = EmergHerbWetlands.5km / TotalPixels.5km,
    CultivCrops.5km.pct = CultivCrops.5km / TotalPixels.5km,
    PerennialIceSnow.5km.pct = PerennialIceSnow.5km / TotalPixels.5km
  )
```

```{r Combining & saving datasets}
## Combining 1-km and 5-km datasets
LC.all <- full_join(LC.1km.final, LC.5km.final, by = c("Site", "Year"))

## Pixel percentages only
# names(LC.all)
LC.pct <- LC.all[,c(1:2,19:33,51:66)]

## Save both as CSV files
write.csv(LC.all, "C:\\Users\\jessicaherrmann\\Desktop\\Delta_NLCDall_1km_5km_2004_2024.csv")
write.csv(LC.pct, "C:\\Users\\jessicaherrmann\\Desktop\\Delta_NLCDpct_1km_5km_2004_2024.csv")
```
