---
title: "DeerCreek_NMDS_AdditionalSiteClusters_JH04.29.2026"
author: "Jessica Herrmann"
date: "2026-04-29"
output: html_document
---
## Background

Updated NMDS script for Deer Creek BMI data analysis - revised to include biological data with rare families removed??? (under 5% per sample event), which was then relativized using the Wisconsin Double Standardization method (relativized by column, then by row/sample). We are further subsetting to sites on DC mainstem (1, 2, 4, 6, 8, 9). Other edit was that Crambidae and Pyralidae were combined under Pyraloidea.

```{r Packages}
library(dplyr)
library(tidyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
require(magrittr)
library(scales)
library(reshape2)
library(vegan)
library(goeveg)
# library(vegan3d)
```

```{r Datasets}
## Previous dataset, for comparison only:
# old <- Rare_removed_BMI_Daily_USGS_Flow_Daily_PRISM_Climate_SOI_PET_Raw_WQ_NA_removed_2003_2022
# old <- old[,c(2:163)]
# names(old)

## Remove NA's from WQ dataset. It usually does not work to run ordinations with NAs in the dataset.
bio <- FINAL_DATASET_BMI_Bio_Flow_Wq_Clim_updated5_2_26
bio <- bio[,c(2:165)]
# head(bio)
nrow(bio) # n = 261 (good) - if n = 265, we still have the duplicate dates error.
names(bio)

names(bio)[66] <- "Pyraloidea" # Correcting spelling error
bio[is.na(bio)] <- 0 # Because vegan doesn't like NA's, we are replacing NA values w/ 0's
# But NOT deleting NA rows

# colSums(is.na(bio)) # Check for NA values
```

```{r Modifying master dataset for ordinations}
## Change Month, Site, WY, Year, Date, Season, and Flow Categories to factors
bio$Site <- as.factor(bio$Site)
bio$Month <- as.factor(bio$Month)
bio$WY <- as.factor(bio$WY)
bio$Year <- as.factor(bio$Year)
bio$Date <- as.factor(bio$Date)
bio$Season <- as.factor(bio$Season)
bio$Cat1 <- as.factor(as.character(bio$Cat1))
bio$Cat3 <- as.factor(as.character(bio$Cat3))
bio$Cat6 <- as.factor(as.character(bio$Cat6))
bio$Cat12 <- as.factor(as.character(bio$Cat12))
## Also make sure all families are numeric values!

## Also set levels for flow categories - will make ordinations easier
bio$Cat1 <- factor(bio$Cat1,levels=c("Dry","Average","Wet"))
bio$Cat3 <- factor(bio$Cat3,levels=c("Dry","Average","Wet"))
bio$Cat6 <- factor(bio$Cat6,levels=c("Dry","Average","Wet"))
bio$Cat12 <- factor(bio$Cat12,levels=c("Dry","Average","Wet"))

## Set colors for each flow category
color.by.cat <- c("Dry" = "coral3", "Average" = "burlywood", "Wet" = "lightskyblue")

## Set levels for season
bio$Season <- factor(bio$Season,levels = c("Base", "Peak"))

## Color code by season
color.by.season <- c("Base" = "cyan3", "Peak" = "darkcyan")


#### UPDATED WATERSHED LOCATIONS TO TEST ####
## Levels by watershed location along mainstem
## Remove non-mainstem sites for this analysis (Squirrel Crk. and Little Deer Crk.)

# 1. Upper DC 1/2 vs. Mid DC 4/6 vs. Lower DC 8/9
# 2. Upper DC 1/2 vs. Mid DC 4 vs. Lower DC 6/8/9
# 3. Upper DC 1/2/4 vs. Mid DC 6 vs. Lower DC 8/9
# 4. Upper DC 1/2/4 vs. Lower 6/8/9
# 5. Upper DC 1/2 vs. Lower 4/6/8/9
# 6. Upper DC 1/2/4/6 vs. Lower 8/9
# 7. Upper DC 1/2/4 vs. Lower DC 8/9
# 8. Upper DC 1/2 vs. Lower 6/8/9

## For all groupings, I'm also using LDC 11/12/13 and SC 15/16.

## Grouping 1
# BIO1 <- bio
# BIO1$Watershed.location <- as.factor(ifelse(bio$Site %in% c(1,2), "Upper DC",
#                           ifelse(bio$Site %in% c(4,6), "Middle DC", 
#                                  ifelse(bio$Site %in% c(8,9), "Lower DC",
#                                         ifelse(bio$Site %in% c(15,16), "Squirrel Creek",
#                                                ifelse(bio$Site %in% c(11,12,13), "Little Deer Creek",
#                                         0))))))
# BIO1 <- BIO1[,c(1:6,165,7:164)]
# BIO1$Watershed.location <- factor(BIO1$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 2
# BIO2 <- bio
# BIO2$Watershed.location <- as.factor(ifelse(BIO2$Site %in% c(1,2), "Upper DC",
#                           ifelse(BIO2$Site %in% c(4), "Middle DC", 
#                                  ifelse(BIO2$Site %in% c(6,8,9), "Lower DC",
#                                         ifelse(BIO2$Site %in% c(15,16), "Squirrel Creek",
#                                                ifelse(BIO2$Site %in% c(11,12,13), "Little Deer Creek",
#                                         0))))))
# BIO2 <- BIO2[,c(1:6,165,7:164)]
# BIO2$Watershed.location <- factor(BIO2$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 3
# BIO3 <- bio
# BIO3$Watershed.location <- as.factor(ifelse(BIO3$Site %in% c(1,2,4), "Upper DC",
#                                  ifelse(BIO3$Site %in% c(6), "Middle DC", 
#                                         ifelse(BIO3$Site %in% c(8,9), "Lower DC",
#                                                ifelse(BIO3$Site %in% c(15,16), "Squirrel Creek",
#                                                       ifelse(BIO3$Site %in% c(11,12,13), "Little Deer Creek",
#                                         0))))))
# BIO3 <- BIO3[,c(1:6,165,7:164)]
# BIO3$Watershed.location <- factor(BIO3$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 4
# BIO4 <- bio
# BIO4$Watershed.location <- as.factor(ifelse(BIO4$Site %in% c(1,2,4), "Upper DC",
#                                      ifelse(BIO4$Site %in% c(6,8,9), "Lower DC",
#                                             ifelse(BIO4$Site %in% c(15,16), "Squirrel Creek",
#                                                    ifelse(BIO4$Site %in% c(11,12,13), "Little Deer Creek",
#                                       0)))))
# BIO4 <- BIO4[,c(1:6,165,7:164)]
# BIO4$Watershed.location <- factor(BIO4$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 5
# BIO5 <- bio
# BIO5$Watershed.location <- as.factor(ifelse(BIO5$Site %in% c(1,2), "Upper DC",
#                                         ifelse(BIO5$Site %in% c(4,6,8,9), "Lower DC",
#                                                ifelse(BIO5$Site %in% c(15,16), "Squirrel Creek",
#                                                       ifelse(BIO5$Site %in% c(11,12,13), "Little Deer Creek",
#                                       0)))))
# BIO5 <- BIO5[,c(1:6,165,7:164)]
# BIO5$Watershed.location <- factor(BIO5$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 6
# BIO6 <- bio
# BIO6$Watershed.location <- as.factor(ifelse(BIO6$Site %in% c(1,2,4,6), "Upper DC",
#                                         ifelse(BIO6$Site %in% c(8,9), "Lower DC",
#                                                ifelse(BIO6$Site %in% c(15,16), "Squirrel Creek",
#                                                       ifelse(BIO6$Site %in% c(11,12,13), "Little Deer Creek",
#                                       0)))))
# BIO6 <- BIO6[,c(1:6,165,7:164)]
# BIO6$Watershed.location <- factor(BIO6$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 7
# BIO7 <- subset(bio, Site != "6")
# BIO7$Watershed.location <- as.factor(ifelse(BIO7$Site %in% c(1,2,4), "Upper DC",
#                                         ifelse(BIO7$Site %in% c(8,9), "Lower DC",
#                                                ifelse(BIO7$Site %in% c(15,16), "Squirrel Creek",
#                                                       ifelse(BIO7$Site %in% c(11,12,13), "Little Deer Creek",
#                                       0)))))
# BIO7 <- BIO7[,c(1:6,165,7:164)]
# BIO7$Watershed.location <- factor(BIO7$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))
# 
# ## Grouping 8
# BIO8 <- subset(bio, Site != "4")
# BIO8$Watershed.location <- as.factor(ifelse(BIO8$Site %in% c(1,2), "Upper DC",
#                                         ifelse(BIO8$Site %in% c(6,8,9), "Lower DC",
#                                                ifelse(BIO8$Site %in% c(15,16), "Squirrel Creek",
#                                                       ifelse(BIO8$Site %in% c(11,12,13), "Little Deer Creek",
#                                       0)))))
# BIO8 <- BIO8[,c(1:6,165,7:164)]
# BIO8$Watershed.location <- factor(BIO8$Watershed.location, levels = c("Upper DC", "Middle DC", "Lower DC", "Squirrel Creek", "Little Deer Creek"))

## Color code by watershed location (for 3 groups and 2 groups)
# color.by.location.3cat <- c("Upper DC" = "blue", "Middle DC" = "dodgerblue", "Lower DC" = "skyblue", "Squirrel Creek" = "olivedrab", "Little Deer Creek" = "bisque3")
# 
# color.by.location.2cat <- c("Upper DC" = "blue", "Lower DC" = "skyblue", "Squirrel Creek" = "olivedrab", "Little Deer Creek" = "bisque3")
```

```{r GROUPING 1 - NMDS Data Prep}
## GROUPINGS 1. Upper DC 1/2 vs. Mid DC 4/6 vs. Lower DC 8/9

#### ALL ####

## METADATA MATRIX
# BIO1.META <- BIO1[,c(1:7)] # Sample meta data (identifiers)
# 
# ## BIOLOGICAL/BMI DATA MATRIX
# BIO1.BMI <- BIO1[,c(8:77)] # Note - BMI numbers must be numeric
# 
# ## ENVIRONMENTAL DATA MATRICES
# BIO1.ENV <- BIO1[,c(1:7, 78:165)] # All environmental variables: meta, bio/WQ/climate, & flow cats
# 
# BIO1.BWC <- BIO1[,c(78:102,106:165)] # Bio, WQ, climate (aka BWC) vars only
# 
# BIO1.BIOVEC <- BIO1[,c(78:102)] # Biological vectors only
# 
# BIO1.CLIM <- BIO1[,c(148:165)] # Climate vectors only
# 
# BIO1.WQ <- BIO1[,c(107:147)] # WQ vectors only
# 
# 
# # CATEGORICAL VARIABLES
# BIO1.flow <- BIO1[,c(103:106)] # Flow categories
# BIO1.season <- BIO1[,c(6)] # Season only
# BIO1.wslocation <- BIO1[,c(7)]






# 1. Upper DC 1/2 vs. Mid DC 4/6 vs. Lower DC 8/9
# 2. Upper DC 1/2 vs. (Mid DC 4) vs. Lower DC 6/8/9    ## Didn't create groupings for indiv. sites 4 & 6 b/c not enough data
# 3. Upper DC 1/2/4 vs. (Mid DC 6) vs. Lower DC 8/9.   ## Didn't create groupings for indiv. sites 4 & 6 b/c not enough data
# 4. Upper DC 1/2/4 vs. Lower 6/8/9
# 5. Upper DC 1/2 vs. Lower 4/6/8/9
# 6. Upper DC 1/2/4/6 vs. Lower 8/9
# 7. Upper DC 1/2/4 vs. Lower DC 8/9
# 8. Upper DC 1/2 vs. Lower 6/8/9


#### UPPER DC POSSIBLE GROUPINGS ####

## GROUPING 1 // UDC 1:
UDC1 <- subset(bio, Site %in% c(1,2))

## METADATA MATRIX
UDC1.META <- UDC1[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
UDC1.BMI <- UDC1[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
UDC1.ENV <- UDC1[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

UDC1.BWC <- UDC1[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

UDC1.BIOVEC <- UDC1[,c(77:101)] # Biological vectors only

UDC1.CLIM <- UDC1[,c(147:164)] # Climate vectors only

UDC1.WQ <- UDC1[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
UDC1.flow <- UDC1[,c(102:105)] # Flow categories
UDC1.season <- UDC1[,c(6)] # Season only



## GROUPING 2 // UDC 2:
UDC2 <- subset(bio, Site %in% c(1,2,4))

## METADATA MATRIX
UDC2.META <- UDC2[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
UDC2.BMI <- UDC2[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
UDC2.ENV <- UDC2[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

UDC2.BWC <- UDC2[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

UDC2.BIOVEC <- UDC2[,c(77:101)] # Biological vectors only

UDC2.CLIM <- UDC2[,c(147:164)] # Climate vectors only

UDC2.WQ <- UDC2[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
UDC2.flow <- UDC2[,c(102:105)] # Flow categories
UDC2.season <- UDC2[,c(6)] # Season only



## GROUPING 3 // UDC 3:
UDC3 <- subset(bio, Site %in% c(1,2,4,6))

## METADATA MATRIX
UDC3.META <- UDC3[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
UDC3.BMI <- UDC3[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
UDC3.ENV <- UDC3[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

UDC3.BWC <- UDC3[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

UDC3.BIOVEC <- UDC3[,c(77:101)] # Biological vectors only

UDC3.CLIM <- UDC3[,c(147:164)] # Climate vectors only

UDC3.WQ <- UDC3[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
UDC3.flow <- UDC3[,c(102:105)] # Flow categories
UDC3.season <- UDC3[,c(6)] # Season only




#### MIDDLE DC POSSIBLE GROUPING(S) ####

## GROUPING 1 // MDC 1
MDC1 <- subset(bio, Site %in% c(4,6))

## METADATA MATRIX
MDC1.META <- MDC1[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
MDC1.BMI <- MDC1[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
MDC1.ENV <- MDC1[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

MDC1.BWC <- MDC1[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

MDC1.BIOVEC <- MDC1[,c(77:101)] # Biological vectors only

MDC1.CLIM <- MDC1[,c(147:164)] # Climate vectors only

MDC1.WQ <- MDC1[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
MDC1.flow <- MDC1[,c(102:105)] # Flow categories
MDC1.season <- MDC1[,c(6)] # Season only





#### LOWER DC POSSIBLE GROUPINGS ####

## GROUPING 1 // LDC 1
LDC1 <- subset(bio, Site %in% c(8,9))

## METADATA MATRIX
LDC1.META <- LDC1[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
LDC1.BMI <- LDC1[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
LDC1.ENV <- LDC1[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

LDC1.BWC <- LDC1[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

LDC1.BIOVEC <- LDC1[,c(77:101)] # Biological vectors only

LDC1.CLIM <- LDC1[,c(147:164)] # Climate vectors only

LDC1.WQ <- LDC1[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
LDC1.flow <- LDC1[,c(102:105)] # Flow categories
LDC1.season <- LDC1[,c(6)] # Season only



## GROUPING 2 // LDC 2
LDC2 <- subset(bio, Site %in% c(6,8,9))

## METADATA MATRIX
LDC2.META <- LDC2[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
LDC2.BMI <- LDC2[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
LDC2.ENV <- LDC2[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

LDC2.BWC <- LDC2[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

LDC2.BIOVEC <- LDC2[,c(77:101)] # Biological vectors only

LDC2.CLIM <- LDC2[,c(147:164)] # Climate vectors only

LDC2.WQ <- LDC2[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
LDC2.flow <- LDC2[,c(102:105)] # Flow categories
LDC2.season <- LDC2[,c(6)] # Season only



## GROUPING 3 // LDC 3
LDC3 <- subset(bio, Site %in% c(4,6,8,9))

## METADATA MATRIX
LDC3.META <- LDC3[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
LDC3.BMI <- LDC3[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
LDC3.ENV <- LDC3[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

LDC3.BWC <- LDC3[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

LDC3.BIOVEC <- LDC3[,c(77:101)] # Biological vectors only

LDC3.CLIM <- LDC3[,c(147:164)] # Climate vectors only

LDC3.WQ <- LDC3[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
LDC3.flow <- LDC3[,c(102:105)] # Flow categories
LDC3.season <- LDC3[,c(6)] # Season only





#### TRIBUTARY GROUPINGS (THESE REMAIN THE SAME THROUGHOUT) ####

## LITTLE DC  // LtDC
LtDC <- subset(bio, Site %in% c(11,12,13))

## METADATA MATRIX
LtDC.META <- LtDC[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
LtDC.BMI <- LtDC[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
LtDC.ENV <- LtDC[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

LtDC.BWC <- LtDC[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

LtDC.BIOVEC <- LtDC[,c(77:101)] # Biological vectors only

LtDC.CLIM <- LtDC[,c(147:164)] # Climate vectors only

LtDC.WQ <- LtDC[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
LtDC.flow <- LtDC[,c(102:105)] # Flow categories
LtDC.season <- LtDC[,c(6)] # Season only



## SQUIRREL CREEK  // SC
SC <- subset(bio, Site %in% c(15,16))

## METADATA MATRIX
SC.META <- SC[,c(1:6)] # Sample meta data (identifiers)

## BIOLOGICAL/BMI DATA MATRIX
SC.BMI <- SC[,c(7:76)] # Note - BMI numbers must be numeric

## ENVIRONMENTAL DATA MATRICES
SC.ENV <- SC[,c(1:7, 76:164)] # All environmental variables: meta, bio/WQ/climate, & flow cats

SC.BWC <- SC[,c(77:101,106:164)] # Bio, WQ, climate (aka BWC) vars only

SC.BIOVEC <- SC[,c(77:101)] # Biological vectors only

SC.CLIM <- SC[,c(147:164)] # Climate vectors only

SC.WQ <- SC[,c(106:146)] # WQ vectors only

# CATEGORICAL VARIABLES
SC.flow <- SC[,c(102:105)] # Flow categories
SC.season <- SC[,c(6)] # Season only
```

```{r UDC 1 - Rare Removed + WDS Ordination}
UDC1.wds <- wisconsin(UDC1.BMI)

# Run the ordination
UDC1.wds.model <- metaMDS(UDC1.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(UDC1.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(UDC1.wds.model)

# Calculate goodness of fit
gof.UDC1 <- goodness(object = UDC1.wds.model)
gof.UDC1

plot(UDC1.wds.model)

# MRPP for WDS dataset, Cat 1
UDC1Cat1.WDS.MRPP <- mrpp(UDC1.wds, UDC1.flow$Cat1, permutations = 999, distance = "bray")
print(UDC1Cat1.WDS.MRPP)

# MRPP, Cat 3
UDC1Cat3.WDS.MRPP <- mrpp(UDC1.wds, UDC1.flow$Cat3, permutations = 999, distance = "bray")
print(UDC1Cat3.WDS.MRPP)

# MRPP, Cat 6
UDC1Cat6.WDS.MRPP <- mrpp(UDC1.wds, UDC1.flow$Cat6, permutations = 999, distance = "bray")
print(UDC1Cat6.WDS.MRPP)

# MRPP, Cat 12
UDC1Cat12.WDS.MRPP <- mrpp(UDC1.wds, UDC1.flow$Cat12, permutations = 999, distance = "bray")
print(UDC1Cat12.WDS.MRPP)
```

``` {r UDC 1 - Significant envrionmental variables}
# # Dimensions 1 & 2 (default)
# # Fit environmental vectors
# UDC1.fit_12.wds <- envfit(UDC1.wds.model, UDC1.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.UDC1.wds <- as.data.frame(UDC1.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.UDC1.wds     <- UDC1.fit_12.wds$vectors$r                             # R-squared
# pvals_12.UDC1.wds  <- UDC1.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC1.sigvec_12.wds <- arrows_12.UDC1.wds %>%
#   mutate(R2 = r2_12.UDC1.wds, pval = pvals_12.UDC1.wds, Variable = rownames(arrows_12.UDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC1.sigvec_12.wds <- UDC1.sigvec_12.wds %>% filter(pval < 0.05)
# 
# 
# # Dimensions 1 & 3
# # Fit environmental vectors
# UDC1.fit_13.wds <- envfit(UDC1.wds.model, UDC1.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.UDC1.wds <- as.data.frame(UDC1.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.UDC1.wds     <- UDC1.fit_13.wds$vectors$r                             # R-squared
# pvals_13.UDC1.wds  <- UDC1.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC1.sigvec_13.wds <- arrows_13.UDC1.wds %>%
#   mutate(R2 = r2_13.UDC1.wds, pval = pvals_13.UDC1.wds, Variable = rownames(arrows_13.UDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC1.sigvec_13.wds <- UDC1.sigvec_13.wds %>% filter(pval < 0.05)
# 
# 
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# UDC1.fit_23.wds <- envfit(UDC1.wds.model, UDC1.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.UDC1.wds <- as.data.frame(UDC1.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.UDC1.wds     <- UDC1.fit_23.wds$vectors$r                             # R-squared
# pvals_23.UDC1.wds  <- UDC1.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC1.sigvec_23.wds <- arrows_23.UDC1.wds %>%
#   mutate(R2 = r2_23.UDC1.wds, pval = pvals_23.UDC1.wds, Variable = rownames(arrows_23.UDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC1.sigvec_23.wds <- UDC1.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r UDC 1 - Flow Cat 1 Model}
# Model with biological data
# plot(UDC1.wds.model, display = "sites", type = "n", main = "UDC 1 Flow Cat 1");   # Creates empty plot
# points(UDC1.wds.model, pch = 21, bg = UDC1$Cat1,                   # Formats points
#        col = "black",cex = 1.5);
# legend("topright", legend = names(color.by.cat),                 # Formats legend
#        fill = color.by.cat, title = "Flow", bty = "n");
# ordiellipse(UDC1.wds.model, groups = UDC1$Cat1,                    # Adds ellipses
#            col = c("coral3", "burlywood", "lightskyblue"),
#            draw = "polygon", label = T, cex = 0.5)
```

```{r UDC 1 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
UDC1.scores <- scores(UDC1.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r UDC 1 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
UDC1.fit1_2.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(1,2))

plot(UDC1.scores[,1], UDC1.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC1.fit1_3.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(1,3))

plot(UDC1.scores[,1], UDC1.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC1.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC1.fit2_3.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(2,3))

plot(UDC1.scores[,2], UDC1.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC1.fit2_3.bio, p.max = 0.05)
```

```{r UDC 1 - Season + Bio Vars}
## DIMENSIONS 1 and 2
# UDC1.fit1_2.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(1,2))
# 
# plot(UDC1.scores[,1], UDC1.scores[,2], # Plots the 1 vs 2 dimensions
#      xlab = "NMDS1", ylab = "NMDS2",
#      pch = 21, bg = color.by.season,  
#      col = "black", cex = 1.5,
#      main = "UDC 1 by Season");
# legend("topright", legend = names(color.by.season),              
#        fill = color.by.season, title = "Flow", bty = "n");
# ordiellipse(UDC1.wds.model, groups = UDC1$UDC1.season,
#             choices = c(1,2), # For ellipses around axes 1 and 2
#             col = c("cyan3","darkcyan"),
#             draw = "polygon", label = T, cex = 0.5);
# plot(UDC1.fit1_2.bio, p.max = 0.05)
```

```{r UDC 1 - Cat 1 Ordination - all sig climate variables}
## DIMENSIONS 1 and 2
UDC1.fit1_2.clim <- envfit(UDC1.wds.model, UDC1.CLIM, permutations = 999, choices = c(1,2))

plot(UDC1.scores[,1], UDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 1v2 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC1.fit1_3.clim <- envfit(UDC1.wds.model, UDC1.CLIM, permutations = 999, choices = c(1,3))

plot(UDC1.scores[,1], UDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 1v3 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC1.fit2_3.clim <- envfit(UDC1.wds.model, UDC1.CLIM, permutations = 999, choices = c(2,3))

plot(UDC1.scores[,2], UDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 2v3 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit2_3.clim, p.max = 0.05)
```

```{r UDC 1 - Cat 1 Ordination - all sig WQ variables}
## DIMENSIONS 1 and 2
UDC1.fit1_2.wq <- envfit(UDC1.wds.model, UDC1.WQ, permutations = 999, choices = c(1,2)) 

plot(UDC1.scores[,1], UDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 1v2 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC1.fit1_3.wq <- envfit(UDC1.wds.model, UDC1.WQ, permutations = 999, choices = c(1,3)) 

plot(UDC1.scores[,1], UDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 1v3 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC1.fit2_3.wq <- envfit(UDC1.wds.model, UDC1.WQ, permutations = 999, choices = c(2,3)) 

plot(UDC1.scores[,2], UDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 1: 1-month Flow\n 2v3 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit2_3.wq, p.max = 0.05)
```
```{r UDC 1 - Cat 3 Bio}
## DIMENSIONS 1 and 2
UDC1.fit1_2.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(1,2))

plot(UDC1.scores[,1], UDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC1.fit1_3.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(1,3))

plot(UDC1.scores[,1], UDC1.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC1.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC1.fit2_3.bio <- envfit(UDC1.wds.model, UDC1.BIOVEC, permutations = 999, choices = c(2,3))

plot(UDC1.scores[,2], UDC1.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC1.fit2_3.bio, p.max = 0.05)
```

```{r UDC 1 - Cat 3 Ordination - all sig climate variables}
## DIMENSIONS 1 and 2
UDC1.fit1_2.clim <- envfit(UDC1.wds.model, UDC1.CLIM, permutations = 999, choices = c(1,2))

plot(UDC1.scores[,1], UDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 1v2 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC1.fit1_3.clim <- envfit(UDC1.wds.model, UDC1.CLIM, permutations = 999, choices = c(1,3))

plot(UDC1.scores[,1], UDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 1v3 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC1.fit2_3.clim <- envfit(UDC1.wds.model, UDC1.CLIM, permutations = 999, choices = c(2,3))

plot(UDC1.scores[,2], UDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 2v3 dimensions \n Sig clim vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit2_3.clim, p.max = 0.05)
```
```{r UDC 1 - Cat 3 Ordination - all sig WQ variables}
## DIMENSIONS 1 and 2
UDC1.fit1_2.wq <- envfit(UDC1.wds.model, UDC1.WQ, permutations = 999, choices = c(1,2)) 

plot(UDC1.scores[,1], UDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC1.fit1_3.wq <- envfit(UDC1.wds.model, UDC1.WQ, permutations = 999, choices = c(1,3)) 

plot(UDC1.scores[,1], UDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC1.fit2_3.wq <- envfit(UDC1.wds.model, UDC1.WQ, permutations = 999, choices = c(2,3)) 

plot(UDC1.scores[,2], UDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 1: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC1.wds.model, groups = UDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC1.fit2_3.wq, p.max = 0.05)
```
```{r UDC 2 - Rare Removed + WDS Ordination}
UDC2.wds <- wisconsin(UDC2.BMI)

# Run the ordination
UDC2.wds.model <- metaMDS(UDC2.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(UDC2.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(UDC2.wds.model)

# Calculate goodness of fit
gof.UDC2 <- goodness(object = UDC2.wds.model)
gof.UDC2

plot(UDC2.wds.model)

# MRPP for WDS dataset, Cat 1
UDC2Cat1.WDS.MRPP <- mrpp(UDC2.wds, UDC2.flow$Cat1, permutations = 999, distance = "bray")
print(UDC2Cat1.WDS.MRPP)

# MRPP, Cat 3
UDC2Cat3.WDS.MRPP <- mrpp(UDC2.wds, UDC2.flow$Cat3, permutations = 999, distance = "bray")
print(UDC2Cat3.WDS.MRPP)

# MRPP, Cat 6
UDC2Cat6.WDS.MRPP <- mrpp(UDC2.wds, UDC2.flow$Cat6, permutations = 999, distance = "bray")
print(UDC2Cat6.WDS.MRPP)

# MRPP, Cat 12
UDC2Cat12.WDS.MRPP <- mrpp(UDC2.wds, UDC2.flow$Cat12, permutations = 999, distance = "bray")
print(UDC2Cat12.WDS.MRPP)
```
``` {r UDC 2 - Significant envrionmental variables}
# Dimensions 1 & 2 (default)
# Fit environmental vectors
# UDC2.fit_12.wds <- envfit(UDC2.wds.model, UDC2.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.UDC2.wds <- as.data.frame(UDC2.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.UDC2.wds     <- UDC2.fit_12.wds$vectors$r                             # R-squared
# pvals_12.UDC2.wds  <- UDC2.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC2.sigvec_12.wds <- arrows_12.UDC2.wds %>%
#   mutate(R2 = r2_12.UDC2.wds, pval = pvals_12.UDC2.wds, Variable = rownames(arrows_12.UDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC2.sigvec_12.wds <- UDC2.sigvec_12.wds %>% filter(pval < 0.05)


# Dimensions 1 & 3
# Fit environmental vectors
# UDC2.fit_13.wds <- envfit(UDC2.wds.model, UDC2.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.UDC2.wds <- as.data.frame(UDC2.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.UDC2.wds     <- UDC2.fit_13.wds$vectors$r                             # R-squared
# pvals_13.UDC2.wds  <- UDC2.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC2.sigvec_13.wds <- arrows_13.UDC2.wds %>%
#   mutate(R2 = r2_13.UDC2.wds, pval = pvals_13.UDC2.wds, Variable = rownames(arrows_13.UDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC2.sigvec_13.wds <- UDC2.sigvec_13.wds %>% filter(pval < 0.05)
# 
# 
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# UDC2.fit_23.wds <- envfit(UDC2.wds.model, UDC2.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.UDC2.wds <- as.data.frame(UDC2.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.UDC2.wds     <- UDC2.fit_23.wds$vectors$r                             # R-squared
# pvals_23.UDC2.wds  <- UDC2.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC2.sigvec_23.wds <- arrows_23.UDC2.wds %>%
#   mutate(R2 = r2_23.UDC2.wds, pval = pvals_23.UDC2.wds, Variable = rownames(arrows_23.UDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC2.sigvec_23.wds <- UDC2.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r UDC 2 - Flow Cat 1 Model}
# Model with biological data
# plot(UDC2.wds.model, display = "sites", type = "n", main = "UDC 2 Flow Cat 1");   # Creates empty plot
# points(UDC2.wds.model, pch = 21, bg = UDC2$Cat1,                   # Formats points
#        col = "black",cex = 1.5);
# legend("topright", legend = names(color.by.cat),                 # Formats legend
#        fill = color.by.cat, title = "Flow", bty = "n");
# ordiellipse(UDC2.wds.model, groups = UDC2$Cat1,                    # Adds ellipses
#            col = c("coral3", "burlywood", "lightskyblue"),
#            draw = "polygon", label = T, cex = 0.5)
```

```{r UDC 2 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
UDC2.scores <- scores(UDC2.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r UDC 2 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
UDC2.fit1_2.bio <- envfit(UDC2.wds.model, UDC2.BIOVEC, permutations = 999, choices = c(1,2))

plot(UDC2.scores[,1], UDC2.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC2.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC2.fit1_3.bio <- envfit(UDC2.wds.model, UDC2.BIOVEC, permutations = 999, choices = c(1,3))

plot(UDC2.scores[,1], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC2.fit2_3.bio <- envfit(UDC2.wds.model, UDC2.BIOVEC, permutations = 999, choices = c(2,3))

plot(UDC2.scores[,2], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit2_3.bio, p.max = 0.05)
```

```{r UDC 2 - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
UDC2.fit1_2.clim <- envfit(UDC2.wds.model, UDC2.CLIM, permutations = 999, choices = c(1,2))

plot(UDC2.scores[,1], UDC2.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC2.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC2.fit1_3.clim <- envfit(UDC2.wds.model, UDC2.CLIM, permutations = 999, choices = c(1,3))

plot(UDC2.scores[,1], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC2.fit2_3.clim <- envfit(UDC2.wds.model, UDC2.CLIM, permutations = 999, choices = c(2,3))

plot(UDC2.scores[,2], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit2_3.clim, p.max = 0.05)
```
```{r UDC 2 - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
UDC2.fit1_2.wq<- envfit(UDC2.wds.model, UDC2.WQ, permutations = 999, choices = c(1,2))

plot(UDC2.scores[,1], UDC2.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC2.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC2.fit1_3.wq <- envfit(UDC2.wds.model, UDC2.WQ, permutations = 999, choices = c(1,3))

plot(UDC2.scores[,1], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC2.fit2_3.wq <- envfit(UDC2.wds.model, UDC2.WQ, permutations = 999, choices = c(2,3))

plot(UDC2.scores[,2], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 2: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit2_3.wq, p.max = 0.05)
```

```{r UDC 2 - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
UDC2.fit1_2.bio <- envfit(UDC2.wds.model, UDC2.BIOVEC, permutations = 999, choices = c(1,2))

plot(UDC2.scores[,1], UDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC2.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC2.fit1_3.bio <- envfit(UDC2.wds.model, UDC2.BIOVEC, permutations = 999, choices = c(1,3))

plot(UDC2.scores[,1], UDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC2.fit2_3.bio <- envfit(UDC2.wds.model, UDC2.BIOVEC, permutations = 999, choices = c(2,3))

plot(UDC2.scores[,2], UDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit2_3.bio, p.max = 0.05)
```

```{r UDC 2 - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
UDC2.fit1_2.clim <- envfit(UDC2.wds.model, UDC2.CLIM, permutations = 999, choices = c(1,2))

plot(UDC2.scores[,1], UDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC2.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC2.fit1_3.clim <- envfit(UDC2.wds.model, UDC2.CLIM, permutations = 999, choices = c(1,3))

plot(UDC2.scores[,1], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC2.fit2_3.clim <- envfit(UDC2.wds.model, UDC2.CLIM, permutations = 999, choices = c(2,3))

plot(UDC2.scores[,2], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit2_3.clim, p.max = 0.05)
```

```{r UDC 2 - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
UDC2.fit1_2.wq<- envfit(UDC2.wds.model, UDC2.WQ, permutations = 999, choices = c(1,2))

plot(UDC2.scores[,1], UDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC2.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC2.fit1_3.wq <- envfit(UDC2.wds.model, UDC2.WQ, permutations = 999, choices = c(1,3))

plot(UDC2.scores[,1], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC2.fit2_3.wq <- envfit(UDC2.wds.model, UDC2.WQ, permutations = 999, choices = c(2,3))

plot(UDC2.scores[,2], UDC2.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 2: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC2.wds.model, groups = UDC2.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC2.fit2_3.wq, p.max = 0.05)
```
```{r UDC 3 - Rare Removed + WDS Ordination}
UDC3.wds <- wisconsin(UDC3.BMI)

# Run the ordination
UDC3.wds.model <- metaMDS(UDC3.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(UDC3.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(UDC3.wds.model)

# Calculate goodness of fit
gof.UDC3 <- goodness(object = UDC3.wds.model)
gof.UDC3

plot(UDC3.wds.model)

# MRPP for WDS dataset, Cat 1
UDC3Cat1.WDS.MRPP <- mrpp(UDC3.wds, UDC3.flow$Cat1, permutations = 999, distance = "bray")
print(UDC3Cat1.WDS.MRPP)

# MRPP, Cat 3
UDC3Cat3.WDS.MRPP <- mrpp(UDC3.wds, UDC3.flow$Cat3, permutations = 999, distance = "bray")
print(UDC3Cat3.WDS.MRPP)

# MRPP, Cat 6
UDC3Cat6.WDS.MRPP <- mrpp(UDC3.wds, UDC3.flow$Cat6, permutations = 999, distance = "bray")
print(UDC3Cat6.WDS.MRPP)

# MRPP, Cat 12
UDC3Cat12.WDS.MRPP <- mrpp(UDC3.wds, UDC3.flow$Cat12, permutations = 999, distance = "bray")
print(UDC3Cat12.WDS.MRPP)
```

``` {r UDC 3 - Significant envrionmental variables}
# Dimensions 1 & 2 (default)
# Fit environmental vectors
# UDC3.fit_12.wds <- envfit(UDC3.wds.model, UDC3.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.UDC3.wds <- as.data.frame(UDC3.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.UDC3.wds     <- UDC3.fit_12.wds$vectors$r                             # R-squared
# pvals_12.UDC3.wds  <- UDC3.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC3.sigvec_12.wds <- arrows_12.UDC3.wds %>%
#   mutate(R2 = r2_12.UDC3.wds, pval = pvals_12.UDC3.wds, Variable = rownames(arrows_12.UDC3.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC3.sigvec_12.wds <- UDC3.sigvec_12.wds %>% filter(pval < 0.05)
#
#
# Dimensions 1 & 3
# Fit environmental vectors
# UDC3.fit_13.wds <- envfit(UDC3.wds.model, UDC3.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.UDC3.wds <- as.data.frame(UDC3.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.UDC3.wds     <- UDC3.fit_13.wds$vectors$r                             # R-squared
# pvals_13.UDC3.wds  <- UDC3.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC3.sigvec_13.wds <- arrows_13.UDC3.wds %>%
#   mutate(R2 = r2_13.UDC3.wds, pval = pvals_13.UDC3.wds, Variable = rownames(arrows_13.UDC3.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC3.sigvec_13.wds <- UDC3.sigvec_13.wds %>% filter(pval < 0.05)
# 
#
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# UDC3.fit_23.wds <- envfit(UDC3.wds.model, UDC3.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.UDC3.wds <- as.data.frame(UDC3.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.UDC3.wds     <- UDC3.fit_23.wds$vectors$r                             # R-squared
# pvals_23.UDC3.wds  <- UDC3.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# UDC3.sigvec_23.wds <- arrows_23.UDC3.wds %>%
#   mutate(R2 = r2_23.UDC3.wds, pval = pvals_23.UDC3.wds, Variable = rownames(arrows_23.UDC3.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# UDC3.sigvec_23.wds <- UDC3.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r UDC 3 - Flow Cat 1 Model}
# Model with biological data
# plot(UDC3.wds.model, display = "sites", type = "n", main = "UDC 3 Flow Cat 1");   # Creates empty plot
# points(UDC3.wds.model, pch = 21, bg = UDC3$Cat1,                   # Formats points
#        col = "black",cex = 1.5);
# legend("topright", legend = names(color.by.cat),                 # Formats legend
#        fill = color.by.cat, title = "Flow", bty = "n");
# ordiellipse(UDC3.wds.model, groups = UDC3$Cat1,                    # Adds ellipses
#            col = c("coral3", "burlywood", "lightskyblue"),
#            draw = "polygon", label = T, cex = 0.5)
```

```{r UDC 3 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
UDC3.scores <- scores(UDC3.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r UDC 3 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
UDC3.fit1_2.bio <- envfit(UDC3.wds.model, UDC3.BIOVEC, permutations = 999, choices = c(1,2))

plot(UDC3.scores[,1], UDC3.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC3.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC3.fit1_3.bio <- envfit(UDC3.wds.model, UDC3.BIOVEC, permutations = 999, choices = c(1,3))

plot(UDC3.scores[,1], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC3.fit2_3.bio <- envfit(UDC3.wds.model, UDC3.BIOVEC, permutations = 999, choices = c(2,3))

plot(UDC3.scores[,2], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit2_3.bio, p.max = 0.05)
```
```{r UDC 3 - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
UDC3.fit1_2.clim <- envfit(UDC3.wds.model, UDC3.CLIM, permutations = 999, choices = c(1,2))

plot(UDC3.scores[,1], UDC3.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC3.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC3.fit1_3.clim <- envfit(UDC3.wds.model, UDC3.CLIM, permutations = 999, choices = c(1,3))

plot(UDC3.scores[,1], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC3.fit2_3.clim <- envfit(UDC3.wds.model, UDC3.CLIM, permutations = 999, choices = c(2,3))

plot(UDC3.scores[,2], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit2_3.clim, p.max = 0.05)
```
```{r UDC 3 - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
UDC3.fit1_2.wq<- envfit(UDC3.wds.model, UDC3.WQ, permutations = 999, choices = c(1,2))

plot(UDC3.scores[,1], UDC3.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC3.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC3.fit1_3.wq <- envfit(UDC3.wds.model, UDC3.WQ, permutations = 999, choices = c(1,3))

plot(UDC3.scores[,1], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC3.fit2_3.wq <- envfit(UDC3.wds.model, UDC3.WQ, permutations = 999, choices = c(2,3))

plot(UDC3.scores[,2], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="UDC 3: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat1,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit2_3.wq, p.max = 0.05)
```
```{r UDC 3 - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
UDC3.fit1_2.bio <- envfit(UDC3.wds.model, UDC3.BIOVEC, permutations = 999, choices = c(1,2))

plot(UDC3.scores[,1], UDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC3.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC3.fit1_3.bio <- envfit(UDC3.wds.model, UDC3.BIOVEC, permutations = 999, choices = c(1,3))

plot(UDC3.scores[,1], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC3.fit2_3.bio <- envfit(UDC3.wds.model, UDC3.BIOVEC, permutations = 999, choices = c(2,3))

plot(UDC3.scores[,2], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit2_3.bio, p.max = 0.05)
```
```{r UDC 3 - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
UDC3.fit1_2.clim <- envfit(UDC3.wds.model, UDC3.CLIM, permutations = 999, choices = c(1,2))

plot(UDC3.scores[,1], UDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC3.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC3.fit1_3.clim <- envfit(UDC3.wds.model, UDC3.CLIM, permutations = 999, choices = c(1,3))

plot(UDC3.scores[,1], UDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC3.fit2_3.clim <- envfit(UDC3.wds.model, UDC3.CLIM, permutations = 999, choices = c(2,3))

plot(UDC3.scores[,2], UDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit2_3.clim, p.max = 0.05)
```
```{r UDC 3 - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
UDC3.fit1_2.wq<- envfit(UDC3.wds.model, UDC3.WQ, permutations = 999, choices = c(1,2))

plot(UDC3.scores[,1], UDC3.scores[,2], # Plots the 1 vs 2 dimensions
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(1,2), # For ellipses around axes 1 and 2
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(UDC3.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
UDC3.fit1_3.wq <- envfit(UDC3.wds.model, UDC3.WQ, permutations = 999, choices = c(1,3))

plot(UDC3.scores[,1], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(1,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
UDC3.fit2_3.wq <- envfit(UDC3.wds.model, UDC3.WQ, permutations = 999, choices = c(2,3))

plot(UDC3.scores[,2], UDC3.scores[,3], # Plots the 1 vs 3 dimensions
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[UDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="UDC 3: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(UDC3.wds.model, groups = UDC3.flow$Cat3,
            choices = c(2,3), # For ellipses around axes 1 and 3
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(UDC3.fit2_3.wq, p.max = 0.05)
```
```{r MDC 1 - Rare Removed + WDS Ordination}
MDC1.wds <- wisconsin(MDC1.BMI)

# Run the ordination
MDC1.wds.model <- metaMDS(MDC1.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(MDC1.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(MDC1.wds.model)

# Calculate goodness of fit
gof.MDC1 <- goodness(object = MDC1.wds.model)
gof.MDC1

plot(MDC1.wds.model)

# MRPP for WDS dataset, Cat 1
MDC1Cat1.WDS.MRPP <- mrpp(MDC1.wds, MDC1.flow$Cat1, permutations = 999, distance = "bray")
print(MDC1Cat1.WDS.MRPP)

# MRPP, Cat 3
MDC1Cat3.WDS.MRPP <- mrpp(MDC1.wds, MDC1.flow$Cat3, permutations = 999, distance = "bray")
print(MDC1Cat3.WDS.MRPP)

# MRPP, Cat 6
MDC1Cat6.WDS.MRPP <- mrpp(MDC1.wds, MDC1.flow$Cat6, permutations = 999, distance = "bray")
print(MDC1Cat6.WDS.MRPP)

# MRPP, Cat 12
MDC1Cat12.WDS.MRPP <- mrpp(MDC1.wds, MDC1.flow$Cat12, permutations = 999, distance = "bray")
print(MDC1Cat12.WDS.MRPP)
```

``` {r MDC 1 - Significant envrionmental variables}
# Dimensions 1 & 2 (default)
# Fit environmental vectors
# MDC1.fit_12.wds <- envfit(MDC1.wds.model, MDC1.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.MDC1.wds <- as.data.frame(MDC1.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.MDC1.wds     <- MDC1.fit_12.wds$vectors$r                             # R-squared
# pvals_12.MDC1.wds  <- MDC1.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# MDC1.sigvec_12.wds <- arrows_12.MDC1.wds %>%
#   mutate(R2 = r2_12.MDC1.wds, pval = pvals_12.MDC1.wds, Variable = rownames(arrows_12.MDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# MDC1.sigvec_12.wds <- MDC1.sigvec_12.wds %>% filter(pval < 0.05)
#
#
# Dimensions 1 & 3
# Fit environmental vectors
# MDC1.fit_13.wds <- envfit(MDC1.wds.model, MDC1.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.MDC1.wds <- as.data.frame(MDC1.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.MDC1.wds     <- MDC1.fit_13.wds$vectors$r                             # R-squared
# pvals_13.MDC1.wds  <- MDC1.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# MDC1.sigvec_13.wds <- arrows_13.MDC1.wds %>%
#   mutate(R2 = r2_13.MDC1.wds, pval = pvals_13.MDC1.wds, Variable = rownames(arrows_13.MDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# MDC1.sigvec_13.wds <- MDC1.sigvec_13.wds %>% filter(pval < 0.05)
# 
#
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# MDC1.fit_23.wds <- envfit(MDC1.wds.model, MDC1.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.MDC1.wds <- as.data.frame(MDC1.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.MDC1.wds     <- MDC1.fit_23.wds$vectors$r                             # R-squared
# pvals_23.MDC1.wds  <- MDC1.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# MDC1.sigvec_23.wds <- arrows_23.MDC1.wds %>%
#   mutate(R2 = r2_23.MDC1.wds, pval = pvals_23.MDC1.wds, Variable = rownames(arrows_23.MDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# MDC1.sigvec_23.wds <- MDC1.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r MDC 1 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
MDC1.scores <- scores(MDC1.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r MDC 1 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
MDC1.fit1_2.bio <- envfit(MDC1.wds.model, MDC1.BIOVEC, permutations = 999, choices = c(1,2))

plot(MDC1.scores[,1], MDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(MDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
MDC1.fit1_3.bio <- envfit(MDC1.wds.model, MDC1.BIOVEC, permutations = 999, choices = c(1,3))

plot(MDC1.scores[,1], MDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
MDC1.fit2_3.bio <- envfit(MDC1.wds.model, MDC1.BIOVEC, permutations = 999, choices = c(2,3))

plot(MDC1.scores[,2], MDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit2_3.bio, p.max = 0.05)
```

```{r MDC 1 - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
MDC1.fit1_2.clim <- envfit(MDC1.wds.model, MDC1.CLIM, permutations = 999, choices = c(1,2))

plot(MDC1.scores[,1], MDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(MDC1.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
MDC1.fit1_3.clim <- envfit(MDC1.wds.model, MDC1.CLIM, permutations = 999, choices = c(1,3))

plot(MDC1.scores[,1], MDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
MDC1.fit2_3.clim <- envfit(MDC1.wds.model, MDC1.CLIM, permutations = 999, choices = c(2,3))

plot(MDC1.scores[,2], MDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit2_3.clim, p.max = 0.05)
```
```{r MDC 1 - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
MDC1.fit1_2.wq <- envfit(MDC1.wds.model, MDC1.WQ, permutations = 999, choices = c(1,2))

plot(MDC1.scores[,1], MDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(MDC1.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
MDC1.fit1_3.wq <- envfit(MDC1.wds.model, MDC1.WQ, permutations = 999, choices = c(1,3))

plot(MDC1.scores[,1], MDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
MDC1.fit2_3.wq <- envfit(MDC1.wds.model, MDC1.WQ, permutations = 999, choices = c(2,3))

plot(MDC1.scores[,2], MDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="MDC 1: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit2_3.wq, p.max = 0.05)
```

```{r MDC 1 - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
MDC1.fit1_2.bio <- envfit(MDC1.wds.model, MDC1.BIOVEC, permutations = 999, choices = c(1,2))

plot(MDC1.scores[,1], MDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(MDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
MDC1.fit1_3.bio <- envfit(MDC1.wds.model, MDC1.BIOVEC, permutations = 999, choices = c(1,3))

plot(MDC1.scores[,1], MDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
MDC1.fit2_3.bio <- envfit(MDC1.wds.model, MDC1.BIOVEC, permutations = 999, choices = c(2,3))

plot(MDC1.scores[,2], MDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit2_3.bio, p.max = 0.05)
```

```{r MDC 1 - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
MDC1.fit1_2.clim <- envfit(MDC1.wds.model, MDC1.CLIM, permutations = 999, choices = c(1,2))

plot(MDC1.scores[,1], MDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(MDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
MDC1.fit1_3.clim <- envfit(MDC1.wds.model, MDC1.CLIM, permutations = 999, choices = c(1,3))

plot(MDC1.scores[,1], MDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
MDC1.fit2_3.clim <- envfit(MDC1.wds.model, MDC1.CLIM, permutations = 999, choices = c(2,3))

plot(MDC1.scores[,2], MDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit2_3.clim, p.max = 0.05)
```

```{r MDC 1 - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
MDC1.fit1_2.wq <- envfit(MDC1.wds.model, MDC1.WQ, permutations = 999, choices = c(1,2))

plot(MDC1.scores[,1], MDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(MDC1.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
MDC1.fit1_3.wq <- envfit(MDC1.wds.model, MDC1.WQ, permutations = 999, choices = c(1,3))

plot(MDC1.scores[,1], MDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
MDC1.fit2_3.wq <- envfit(MDC1.wds.model, MDC1.WQ, permutations = 999, choices = c(2,3))

plot(MDC1.scores[,2], MDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[MDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="MDC 1: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(MDC1.wds.model, groups = MDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(MDC1.fit2_3.wq, p.max = 0.05)
```

```{r LDC 1 - Rare Removed + WDS Ordination}
LDC1.wds <- wisconsin(LDC1.BMI)

# Run the ordination
LDC1.wds.model <- metaMDS(LDC1.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(LDC1.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(LDC1.wds.model)

# Calculate goodness of fit
gof.LDC1 <- goodness(object = LDC1.wds.model)
gof.LDC1

plot(LDC1.wds.model)

# MRPP for WDS dataset, Cat 1
LDC1Cat1.WDS.MRPP <- mrpp(LDC1.wds, LDC1.flow$Cat1, permutations = 999, distance = "bray")
print(LDC1Cat1.WDS.MRPP)

# MRPP, Cat 3
LDC1Cat3.WDS.MRPP <- mrpp(LDC1.wds, LDC1.flow$Cat3, permutations = 999, distance = "bray")
print(LDC1Cat3.WDS.MRPP)

# MRPP, Cat 6
LDC1Cat6.WDS.MRPP <- mrpp(LDC1.wds, LDC1.flow$Cat6, permutations = 999, distance = "bray")
print(LDC1Cat6.WDS.MRPP)

# MRPP, Cat 12
LDC1Cat12.WDS.MRPP <- mrpp(LDC1.wds, LDC1.flow$Cat12, permutations = 999, distance = "bray")
print(LDC1Cat12.WDS.MRPP)
```

``` {r LCD 1 - Significant envrionmental variables}
# Dimensions 1 & 2 (default)
# Fit environmental vectors
# LDC1.fit_12.wds <- envfit(LDC1.wds.model, LDC1.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.LDC1.wds <- as.data.frame(LDC1.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.LDC1.wds     <- LDC1.fit_12.wds$vectors$r                             # R-squared
# pvals_12.LDC1.wds  <- LDC1.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC1.sigvec_12.wds <- arrows_12.LDC1.wds %>%
#   mutate(R2 = r2_12.LDC1.wds, pval = pvals_12.LDC1.wds, Variable = rownames(arrows_12.LDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC1.sigvec_12.wds <- LDC1.sigvec_12.wds %>% filter(pval < 0.05)
#
#
# Dimensions 1 & 3
# Fit environmental vectors
# LDC1.fit_13.wds <- envfit(LDC1.wds.model, LDC1.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.LDC1.wds <- as.data.frame(LDC1.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.LDC1.wds     <- LDC1.fit_13.wds$vectors$r                             # R-squared
# pvals_13.LDC1.wds  <- LDC1.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC1.sigvec_13.wds <- arrows_13.LDC1.wds %>%
#   mutate(R2 = r2_13.LDC1.wds, pval = pvals_13.LDC1.wds, Variable = rownames(arrows_13.LDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC1.sigvec_13.wds <- LDC1.sigvec_13.wds %>% filter(pval < 0.05)
# 
#
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# LDC1.fit_23.wds <- envfit(LDC1.wds.model, LDC1.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.LDC1.wds <- as.data.frame(LDC1.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.LDC1.wds     <- LDC1.fit_23.wds$vectors$r                             # R-squared
# pvals_23.LDC1.wds  <- LDC1.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC1.sigvec_23.wds <- arrows_23.LDC1.wds %>%
#   mutate(R2 = r2_23.LDC1.wds, pval = pvals_23.LDC1.wds, Variable = rownames(arrows_23.LDC1.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC1.sigvec_23.wds <- LDC1.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r LDC 1 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
LDC1.scores <- scores(LDC1.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r LDC 1 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
LDC1.fit1_2.bio <- envfit(LDC1.wds.model, LDC1.BIOVEC, permutations = 999, choices = c(1,2))

plot(LDC1.scores[,1], LDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC1.fit1_3.bio <- envfit(LDC1.wds.model, LDC1.BIOVEC, permutations = 999, choices = c(1,3))

plot(LDC1.scores[,1], LDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC1.fit2_3.bio <- envfit(LDC1.wds.model, LDC1.BIOVEC, permutations = 999, choices = c(2,3))

plot(LDC1.scores[,2], LDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit2_3.bio, p.max = 0.05)
```

```{r LDC 1 - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
LDC1.fit1_2.clim <- envfit(LDC1.wds.model, LDC1.CLIM, permutations = 999, choices = c(1,2))

plot(LDC1.scores[,1], LDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC1.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC1.fit1_3.clim <- envfit(LDC1.wds.model, LDC1.CLIM, permutations = 999, choices = c(1,3))

plot(LDC1.scores[,1], LDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC1.fit2_3.clim <- envfit(LDC1.wds.model, LDC1.CLIM, permutations = 999, choices = c(2,3))

plot(LDC1.scores[,2], LDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit2_3.clim, p.max = 0.05)
```

```{r LDC 1 - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
LDC1.fit1_2.wq <- envfit(LDC1.wds.model, LDC1.WQ, permutations = 999, choices = c(1,2))

plot(LDC1.scores[,1], LDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC1.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC1.fit1_3.wq <- envfit(LDC1.wds.model, LDC1.WQ, permutations = 999, choices = c(1,3))

plot(LDC1.scores[,1], LDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC1.fit2_3.wq <- envfit(LDC1.wds.model, LDC1.WQ, permutations = 999, choices = c(2,3))

plot(LDC1.scores[,2], LDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 1: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit2_3.wq, p.max = 0.05)
```

```{r LDC 1 - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
LDC1.fit1_2.bio <- envfit(LDC1.wds.model, LDC1.BIOVEC, permutations = 999, choices = c(1,2))

plot(LDC1.scores[,1], LDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC1.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC1.fit1_3.bio <- envfit(LDC1.wds.model, LDC1.BIOVEC, permutations = 999, choices = c(1,3))

plot(LDC1.scores[,1], LDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC1.fit2_3.bio <- envfit(LDC1.wds.model, LDC1.BIOVEC, permutations = 999, choices = c(2,3))

plot(LDC1.scores[,2], LDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit2_3.bio, p.max = 0.05)
```
```{r LDC 1 - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
LDC1.fit1_2.clim <- envfit(LDC1.wds.model, LDC1.CLIM, permutations = 999, choices = c(1,2))

plot(LDC1.scores[,1], LDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC1.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC1.fit1_3.clim <- envfit(LDC1.wds.model, LDC1.CLIM, permutations = 999, choices = c(1,3))

plot(LDC1.scores[,1], LDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC1.fit2_3.clim <- envfit(LDC1.wds.model, LDC1.CLIM, permutations = 999, choices = c(2,3))

plot(LDC1.scores[,2], LDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit2_3.clim, p.max = 0.05)
```
```{r LDC 1 - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
LDC1.fit1_2.wq <- envfit(LDC1.wds.model, LDC1.WQ, permutations = 999, choices = c(1,2))

plot(LDC1.scores[,1], LDC1.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC1.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC1.fit1_3.wq <- envfit(LDC1.wds.model, LDC1.WQ, permutations = 999, choices = c(1,3))

plot(LDC1.scores[,1], LDC1.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC1.fit2_3.wq <- envfit(LDC1.wds.model, LDC1.WQ, permutations = 999, choices = c(2,3))

plot(LDC1.scores[,2], LDC1.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC1.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 1: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC1.wds.model, groups = LDC1.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC1.fit2_3.wq, p.max = 0.05)
```

```{r LDC 2 - Rare Removed + WDS Ordination}
LDC2.wds <- wisconsin(LDC2.BMI)

# Run the ordination
LDC2.wds.model <- metaMDS(LDC2.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(LDC2.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(LDC2.wds.model)

# Calculate goodness of fit
gof.LDC2 <- goodness(object = LDC2.wds.model)
gof.LDC2

plot(LDC2.wds.model)

# MRPP for WDS dataset, Cat 1
LDC2Cat1.WDS.MRPP <- mrpp(LDC2.wds, LDC2.flow$Cat1, permutations = 999, distance = "bray")
print(LDC2Cat1.WDS.MRPP)

# MRPP, Cat 3
LDC2Cat3.WDS.MRPP <- mrpp(LDC2.wds, LDC2.flow$Cat3, permutations = 999, distance = "bray")
print(LDC2Cat3.WDS.MRPP)

# MRPP, Cat 6
LDC2Cat6.WDS.MRPP <- mrpp(LDC2.wds, LDC2.flow$Cat6, permutations = 999, distance = "bray")
print(LDC2Cat6.WDS.MRPP)

# MRPP, Cat 12
LDC2Cat12.WDS.MRPP <- mrpp(LDC2.wds, LDC2.flow$Cat12, permutations = 999, distance = "bray")
print(LDC2Cat12.WDS.MRPP)
```

``` {r LCD 2 - Significant envrionmental variables}
# Dimensions 1 & 2 (default)
# Fit environmental vectors
# LDC2.fit_12.wds <- envfit(LDC2.wds.model, LDC2.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.LDC2.wds <- as.data.frame(LDC2.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.LDC2.wds     <- LDC2.fit_12.wds$vectors$r                             # R-squared
# pvals_12.LDC2.wds  <- LDC2.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC2.sigvec_12.wds <- arrows_12.LDC2.wds %>%
#   mutate(R2 = r2_12.LDC2.wds, pval = pvals_12.LDC2.wds, Variable = rownames(arrows_12.LDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC2.sigvec_12.wds <- LDC2.sigvec_12.wds %>% filter(pval < 0.05)
#
#
# Dimensions 1 & 3
# Fit environmental vectors
# LDC2.fit_13.wds <- envfit(LDC2.wds.model, LDC2.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.LDC2.wds <- as.data.frame(LDC2.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.LDC2.wds     <- LDC2.fit_13.wds$vectors$r                             # R-squared
# pvals_13.LDC2.wds  <- LDC2.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC2.sigvec_13.wds <- arrows_13.LDC2.wds %>%
#   mutate(R2 = r2_13.LDC2.wds, pval = pvals_13.LDC2.wds, Variable = rownames(arrows_13.LDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC2.sigvec_13.wds <- LDC2.sigvec_13.wds %>% filter(pval < 0.05)
# 
#
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# LDC2.fit_23.wds <- envfit(LDC2.wds.model, LDC2.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.LDC2.wds <- as.data.frame(LDC2.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.LDC2.wds     <- LDC2.fit_23.wds$vectors$r                             # R-squared
# pvals_23.LDC2.wds  <- LDC2.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC2.sigvec_23.wds <- arrows_23.LDC2.wds %>%
#   mutate(R2 = r2_23.LDC2.wds, pval = pvals_23.LDC2.wds, Variable = rownames(arrows_23.LDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC2.sigvec_23.wds <- LDC2.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r LDC 2 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
LDC2.scores <- scores(LDC2.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r LDC 2 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
LDC2.fit1_2.bio <- envfit(LDC2.wds.model, LDC2.BIOVEC, permutations = 999, choices = c(1,2))

plot(LDC2.scores[,1], LDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC2.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC2.fit1_3.bio <- envfit(LDC2.wds.model, LDC2.BIOVEC, permutations = 999, choices = c(1,3))

plot(LDC2.scores[,1], LDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC2.fit2_3.bio <- envfit(LDC2.wds.model, LDC2.BIOVEC, permutations = 999, choices = c(2,3))

plot(LDC2.scores[,2], LDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit2_3.bio, p.max = 0.05)
```

```{r LDC 2 - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
LDC2.fit1_2.clim <- envfit(LDC2.wds.model, LDC2.CLIM, permutations = 999, choices = c(1,2))

plot(LDC2.scores[,1], LDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC2.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC2.fit1_3.clim <- envfit(LDC2.wds.model, LDC2.CLIM, permutations = 999, choices = c(1,3))

plot(LDC2.scores[,1], LDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC2.fit2_3.clim <- envfit(LDC2.wds.model, LDC2.CLIM, permutations = 999, choices = c(2,3))

plot(LDC2.scores[,2], LDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit2_3.clim, p.max = 0.05)
```

```{r LDC 2 - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
LDC2.fit1_2.wq <- envfit(LDC2.wds.model, LDC2.WQ, permutations = 999, choices = c(1,2))

plot(LDC2.scores[,1], LDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC2.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC2.fit1_3.wq <- envfit(LDC2.wds.model, LDC2.WQ, permutations = 999, choices = c(1,3))

plot(LDC2.scores[,1], LDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC2.fit2_3.wq <- envfit(LDC2.wds.model, LDC2.WQ, permutations = 999, choices = c(2,3))

plot(LDC2.scores[,2], LDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 2: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit2_3.wq, p.max = 0.05)
```

```{r LDC 2 - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
LDC2.fit1_2.bio <- envfit(LDC2.wds.model, LDC2.BIOVEC, permutations = 999, choices = c(1,2))

plot(LDC2.scores[,1], LDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC2.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC2.fit1_3.bio <- envfit(LDC2.wds.model, LDC2.BIOVEC, permutations = 999, choices = c(1,3))

plot(LDC2.scores[,1], LDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC2.fit2_3.bio <- envfit(LDC2.wds.model, LDC2.BIOVEC, permutations = 999, choices = c(2,3))

plot(LDC2.scores[,2], LDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit2_3.bio, p.max = 0.05)
```

```{r LDC 2 - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
LDC2.fit1_2.clim <- envfit(LDC2.wds.model, LDC2.CLIM, permutations = 999, choices = c(1,2))

plot(LDC2.scores[,1], LDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC2.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC2.fit1_3.clim <- envfit(LDC2.wds.model, LDC2.CLIM, permutations = 999, choices = c(1,3))

plot(LDC2.scores[,1], LDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC2.fit2_3.clim <- envfit(LDC2.wds.model, LDC2.CLIM, permutations = 999, choices = c(2,3))

plot(LDC2.scores[,2], LDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit2_3.clim, p.max = 0.05)
```

```{r LDC 2 - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
LDC2.fit1_2.wq <- envfit(LDC2.wds.model, LDC2.WQ, permutations = 999, choices = c(1,2))

plot(LDC2.scores[,1], LDC2.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC2.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC2.fit1_3.wq <- envfit(LDC2.wds.model, LDC2.WQ, permutations = 999, choices = c(1,3))

plot(LDC2.scores[,1], LDC2.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC2.fit2_3.wq <- envfit(LDC2.wds.model, LDC2.WQ, permutations = 999, choices = c(2,3))

plot(LDC2.scores[,2], LDC2.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC2.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 2: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC2.wds.model, groups = LDC2.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC2.fit2_3.wq, p.max = 0.05)
```

```{r LDC 3 - Rare Removed + WDS Ordination}
LDC3.wds <- wisconsin(LDC3.BMI)

# Run the ordination
LDC3.wds.model <- metaMDS(LDC3.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(LDC3.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(LDC3.wds.model)

# Calculate goodness of fit
gof.LDC3 <- goodness(object = LDC3.wds.model)
gof.LDC3

plot(LDC3.wds.model)

# MRPP for WDS dataset, Cat 1
LDC3Cat1.WDS.MRPP <- mrpp(LDC3.wds, LDC3.flow$Cat1, permutations = 999, distance = "bray")
print(LDC3Cat1.WDS.MRPP)

# MRPP, Cat 3
LDC3Cat3.WDS.MRPP <- mrpp(LDC3.wds, LDC3.flow$Cat3, permutations = 999, distance = "bray")
print(LDC3Cat3.WDS.MRPP)

# MRPP, Cat 6
LDC3Cat6.WDS.MRPP <- mrpp(LDC3.wds, LDC3.flow$Cat6, permutations = 999, distance = "bray")
print(LDC3Cat6.WDS.MRPP)

# MRPP, Cat 12
LDC3Cat12.WDS.MRPP <- mrpp(LDC3.wds, LDC3.flow$Cat12, permutations = 999, distance = "bray")
print(LDC3Cat12.WDS.MRPP)
```

``` {r LCD 2 - Significant envrionmental variables}
# Dimensions 1 & 2 (default)
# Fit environmental vectors
# LDC2.fit_12.wds <- envfit(LDC2.wds.model, LDC2.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.LDC2.wds <- as.data.frame(LDC2.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.LDC2.wds     <- LDC2.fit_12.wds$vectors$r                             # R-squared
# pvals_12.LDC2.wds  <- LDC2.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC2.sigvec_12.wds <- arrows_12.LDC2.wds %>%
#   mutate(R2 = r2_12.LDC2.wds, pval = pvals_12.LDC2.wds, Variable = rownames(arrows_12.LDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC2.sigvec_12.wds <- LDC2.sigvec_12.wds %>% filter(pval < 0.05)
#
#
# Dimensions 1 & 3
# Fit environmental vectors
# LDC2.fit_13.wds <- envfit(LDC2.wds.model, LDC2.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.LDC2.wds <- as.data.frame(LDC2.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.LDC2.wds     <- LDC2.fit_13.wds$vectors$r                             # R-squared
# pvals_13.LDC2.wds  <- LDC2.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC2.sigvec_13.wds <- arrows_13.LDC2.wds %>%
#   mutate(R2 = r2_13.LDC2.wds, pval = pvals_13.LDC2.wds, Variable = rownames(arrows_13.LDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC2.sigvec_13.wds <- LDC2.sigvec_13.wds %>% filter(pval < 0.05)
# 
#
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# LDC2.fit_23.wds <- envfit(LDC2.wds.model, LDC2.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.LDC2.wds <- as.data.frame(LDC2.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.LDC2.wds     <- LDC2.fit_23.wds$vectors$r                             # R-squared
# pvals_23.LDC2.wds  <- LDC2.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LDC2.sigvec_23.wds <- arrows_23.LDC2.wds %>%
#   mutate(R2 = r2_23.LDC2.wds, pval = pvals_23.LDC2.wds, Variable = rownames(arrows_23.LDC2.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LDC2.sigvec_23.wds <- LDC2.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r LDC 3 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
LDC3.scores <- scores(LDC3.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r LDC 3 - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
LDC3.fit1_2.bio <- envfit(LDC3.wds.model, LDC3.BIOVEC, permutations = 999, choices = c(1,2))

plot(LDC3.scores[,1], LDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC3.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC3.fit1_3.bio <- envfit(LDC3.wds.model, LDC3.BIOVEC, permutations = 999, choices = c(1,3))

plot(LDC3.scores[,1], LDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC3.fit2_3.bio <- envfit(LDC3.wds.model, LDC3.BIOVEC, permutations = 999, choices = c(2,3))

plot(LDC3.scores[,2], LDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit2_3.bio, p.max = 0.05)
```

```{r LDC 3 - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
LDC3.fit1_2.clim <- envfit(LDC3.wds.model, LDC3.CLIM, permutations = 999, choices = c(1,2))

plot(LDC3.scores[,1], LDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC3.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC3.fit1_3.clim <- envfit(LDC3.wds.model, LDC3.CLIM, permutations = 999, choices = c(1,3))

plot(LDC3.scores[,1], LDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC3.fit2_3.clim <- envfit(LDC3.wds.model, LDC3.CLIM, permutations = 999, choices = c(2,3))

plot(LDC3.scores[,2], LDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit2_3.clim, p.max = 0.05)
```

```{r LDC 3 - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
LDC3.fit1_2.wq <- envfit(LDC3.wds.model, LDC3.WQ, permutations = 999, choices = c(1,2))

plot(LDC3.scores[,1], LDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC3.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC3.fit1_3.wq <- envfit(LDC3.wds.model, LDC3.WQ, permutŵtions = 999, choices = c(1,3))

plot(LDC3.scores[,1], LDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC3.fit2_3.wq <- envfit(LDC3.wds.model, LDC3.WQ, permutations = 999, choices = c(2,3))

plot(LDC3.scores[,2], LDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LDC 3: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit2_3.wq, p.max = 0.05)
```

```{r LDC 3 - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
LDC3.fit1_2.bio <- envfit(LDC3.wds.model, LDC3.BIOVEC, permutations = 999, choices = c(1,2))

plot(LDC3.scores[,1], LDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC3.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC3.fit1_3.bio <- envfit(LDC3.wds.model, LDC3.BIOVEC, permutations = 999, choices = c(1,3))

plot(LDC3.scores[,1], LDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC3.fit2_3.bio <- envfit(LDC3.wds.model, LDC3.BIOVEC, permutations = 999, choices = c(2,3))

plot(LDC3.scores[,2], LDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit2_3.bio, p.max = 0.05)
```

```{r LDC 3 - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
LDC3.fit1_2.clim <- envfit(LDC3.wds.model, LDC3.CLIM, permutations = 999, choices = c(1,2))

plot(LDC3.scores[,1], LDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC3.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC3.fit1_3.clim <- envfit(LDC3.wds.model, LDC3.CLIM, permutations = 999, choices = c(1,3))

plot(LDC3.scores[,1], LDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC3.fit2_3.clim <- envfit(LDC3.wds.model, LDC3.CLIM, permutations = 999, choices = c(2,3))

plot(LDC3.scores[,2], LDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit2_3.clim, p.max = 0.05)
```

```{r LDC 3 - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
LDC3.fit1_2.wq <- envfit(LDC3.wds.model, LDC3.WQ, permutations = 999, choices = c(1,2))

plot(LDC3.scores[,1], LDC3.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LDC3.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LDC3.fit1_3.wq <- envfit(LDC3.wds.model, LDC3.WQ, permutŵtions = 999, choices = c(1,3))

plot(LDC3.scores[,1], LDC3.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LDC3.fit2_3.wq <- envfit(LDC3.wds.model, LDC3.WQ, permutations = 999, choices = c(2,3))

plot(LDC3.scores[,2], LDC3.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LDC3.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LDC 3: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LDC3.wds.model, groups = LDC3.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LDC3.fit2_3.wq, p.max = 0.05)
```
```{r LtDC - Rare Removed + WDS Ordination}
LtDC.wds <- wisconsin(LtDC.BMI)

# Run the ordination
LtDC.wds.model <- metaMDS(LtDC.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(LtDC.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(LtDC.wds.model)

# Calculate goodness of fit
gof.LtDC <- goodness(object = LtDC.wds.model)
gof.LtDC

plot(LtDC.wds.model)

# MRPP for WDS dataset, Cat 1
LtDCCat1.WDS.MRPP <- mrpp(LtDC.wds, LtDC.flow$Cat1, permutations = 999, distance = "bray")
print(LtDCCat1.WDS.MRPP)

# MRPP, Cat 3
LtDCCat3.WDS.MRPP <- mrpp(LtDC.wds, LtDC.flow$Cat3, permutations = 999, distance = "bray")
print(LtDCCat3.WDS.MRPP)

# MRPP, Cat 6
LtDCCat6.WDS.MRPP <- mrpp(LtDC.wds, LtDC.flow$Cat6, permutations = 999, distance = "bray")
print(LtDCCat6.WDS.MRPP)

# MRPP, Cat 12
LtDCCat12.WDS.MRPP <- mrpp(LtDC.wds, LtDC.flow$Cat12, permutations = 999, distance = "bray")
print(LtDCCat12.WDS.MRPP)
```

``` {r LtDC - Significant envrionmental variables}
# # Dimensions 1 & 2 (default)
# # Fit environmental vectors
# LtDC.fit_12.wds <- envfit(LtDC.wds.model, LtDC.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.LtDC.wds <- as.data.frame(LtDC.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.LtDC.wds     <- LtDC.fit_12.wds$vectors$r                             # R-squared
# pvals_12.LtDC.wds  <- LtDC.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LtDC.sigvec_12.wds <- arrows_12.LtDC.wds %>%
#   mutate(R2 = r2_12.LtDC.wds, pval = pvals_12.LtDC.wds, Variable = rownames(arrows_12.LtDC.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LtDC.sigvec_12.wds <- LtDC.sigvec_12.wds %>% filter(pval < 0.05)
# 
# 
# Dimensions 1 & 3
# Fit environmental vectors
# LtDC.fit_13.wds <- envfit(LtDC.wds.model, LtDC.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.LtDC.wds <- as.data.frame(LtDC.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.LtDC.wds     <- LtDC.fit_13.wds$vectors$r                             # R-squared
# pvals_13.LtDC.wds  <- LtDC.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LtDC.sigvec_13.wds <- arrows_13.LtDC.wds %>%
#   mutate(R2 = r2_13.LtDC.wds, pval = pvals_13.LtDC.wds, Variable = rownames(arrows_13.LtDC.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LtDC.sigvec_13.wds <- LtDC.sigvec_13.wds %>% filter(pval < 0.05)
# 
# 
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# LtDC.fit_23.wds <- envfit(LtDC.wds.model, LtDC.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.LtDC.wds <- as.data.frame(LtDC.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.LtDC.wds     <- LtDC.fit_23.wds$vectors$r                             # R-squared
# pvals_23.LtDC.wds  <- LtDC.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# LtDC.sigvec_23.wds <- arrows_23.LtDC.wds %>%
#   mutate(R2 = r2_23.LtDC.wds, pval = pvals_23.LtDC.wds, Variable = rownames(arrows_23.LtDC.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# LtDC.sigvec_23.wds <- LtDC.sigvec_23.wds %>% filter(pval < 0.05)
```

```{r LDC 1 - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
LtDC.scores <- scores(LtDC.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r LtDC - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
LtDC.fit1_2.bio <- envfit(LtDC.wds.model, LtDC.BIOVEC, permutations = 999, choices = c(1,2))

plot(LtDC.scores[,1], LtDC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LtDC.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LtDC.fit1_3.bio <- envfit(LtDC.wds.model, LtDC.BIOVEC, permutations = 999, choices = c(1,3))

plot(LtDC.scores[,1], LtDC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LtDC.fit2_3.bio <- envfit(LtDC.wds.model, LtDC.BIOVEC, permutations = 999, choices = c(2,3))

plot(LtDC.scores[,2], LtDC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit2_3.bio, p.max = 0.05)
```

```{r LtDC - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
LtDC.fit1_2.clim <- envfit(LtDC.wds.model, LtDC.CLIM, permutations = 999, choices = c(1,2))

plot(LtDC.scores[,1], LtDC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LtDC.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LtDC.fit1_3.clim <- envfit(LtDC.wds.model, LtDC.CLIM, permutations = 999, choices = c(1,3))

plot(LtDC.scores[,1], LtDC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LtDC.fit2_3.clim <- envfit(LtDC.wds.model, LtDC.CLIM, permutations = 999, choices = c(2,3))

plot(LtDC.scores[,2], LtDC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit2_3.clim, p.max = 0.05)
```

```{r LtDC - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
LtDC.fit1_2.wq <- envfit(LtDC.wds.model, LtDC.WQ, permutations = 999, choices = c(1,2))

plot(LtDC.scores[,1], LtDC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LtDC.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LtDC.fit1_3.wq <- envfit(LtDC.wds.model, LtDC.WQ, permutations = 999, choices = c(1,3))

plot(LtDC.scores[,1], LtDC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 1v3 dimensions \n Sig WQw vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LtDC.fit2_3.wq <- envfit(LtDC.wds.model, LtDC.WQ, permutations = 999, choices = c(2,3))

plot(LtDC.scores[,2], LtDC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="LtDC: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit2_3.wq, p.max = 0.05)
```

```{r LtDC - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
LtDC.fit1_2.bio <- envfit(LtDC.wds.model, LtDC.BIOVEC, permutations = 999, choices = c(1,2))

plot(LtDC.scores[,1], LtDC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LtDC.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
LtDC.fit1_3.bio <- envfit(LtDC.wds.model, LtDC.BIOVEC, permutations = 999, choices = c(1,3))

plot(LtDC.scores[,1], LtDC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
LtDC.fit2_3.bio <- envfit(LtDC.wds.model, LtDC.BIOVEC, permutations = 999, choices = c(2,3))

plot(LtDC.scores[,2], LtDC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit2_3.bio, p.max = 0.05)
```

```{r LtDC - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
LtDC.fit1_2.clim <- envfit(LtDC.wds.model, LtDC.CLIM, permutations = 999, choices = c(1,2))

plot(LtDC.scores[,1], LtDC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LtDC.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
LtDC.fit1_3.clim <- envfit(LtDC.wds.model, LtDC.CLIM, permutations = 999, choices = c(1,3))

plot(LtDC.scores[,1], LtDC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 1v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
LtDC.fit2_3.clim <- envfit(LtDC.wds.model, LtDC.CLIM, permutations = 999, choices = c(2,3))

plot(LtDC.scores[,2], LtDC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit2_3.clim, p.max = 0.05)
```
```{r LtDC - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
LtDC.fit1_2.wq <- envfit(LtDC.wds.model, LtDC.WQ, permutations = 999, choices = c(1,2))

plot(LtDC.scores[,1], LtDC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(LtDC.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
LtDC.fit1_3.wq <- envfit(LtDC.wds.model, LtDC.WQ, permutations = 999, choices = c(1,3))

plot(LtDC.scores[,1], LtDC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 1v3 dimensions \n Sig WQw vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
LtDC.fit2_3.wq <- envfit(LtDC.wds.model, LtDC.WQ, permutations = 999, choices = c(2,3))

plot(LtDC.scores[,2], LtDC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[LtDC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="LtDC: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(LtDC.wds.model, groups = LtDC.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(LtDC.fit2_3.wq, p.max = 0.05)
```
```{r SC - Rare Removed + WDS Ordination}
SC.wds <- wisconsin(SC.BMI)

# Run the ordination
SC.wds.model <- metaMDS(SC.wds, distance = "bray", 
                                  autotransform = FALSE,
                                  k = 3,
                                  maxit = 999,
                                  trymax = 500,
                                  wascores = TRUE)

screeplot_NMDS(SC.wds, distance = "bray", k = 6, trymax = 500, autotransform = FALSE)
# Lower stress value indicates better fit

# Shepard plot
stressplot(SC.wds.model)

# Calculate goodness of fit
gof.SC <- goodness(object = SC.wds.model)
gof.SC

plot(SC.wds.model)

# MRPP for WDS dataset, Cat 1
SCCat1.WDS.MRPP <- mrpp(SC.wds, SC.flow$Cat1, permutations = 999, distance = "bray")
print(SCCat1.WDS.MRPP)

# MRPP, Cat 3
SCCat3.WDS.MRPP <- mrpp(SC.wds, SC.flow$Cat3, permutations = 999, distance = "bray")
print(SCCat3.WDS.MRPP)

# MRPP, Cat 6
SCCat6.WDS.MRPP <- mrpp(SC.wds, SC.flow$Cat6, permutations = 999, distance = "bray")
print(SCCat6.WDS.MRPP)

# MRPP, Cat 12
SCCat12.WDS.MRPP <- mrpp(SC.wds, SC.flow$Cat12, permutations = 999, distance = "bray")
print(SCCat12.WDS.MRPP)
```

``` {r SC - Significant envrionmental variables}
# # Dimensions 1 & 2 (default)
# # Fit environmental vectors
# SCC.fit_12.wds <- envfit(SC.wds.model, SC.ENV, permutations = 999, choices = c(1,2))
# 
# # Extract vector info
# arrows_12.SC.wds <- as.data.frame(SC.fit_12.wds$vectors$arrows)         # NMDS axis scores
# r2_12.SC.wds     <- SC.fit_12.wds$vectors$r                             # R-squared
# pvals_12.SC.wds  <- SC.fit_12.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# SC.sigvec_12.wds <- arrows_12.SC.wds %>%
#   mutate(R2 = r2_12.SC.wds, pval = pvals_12.SC.wds, Variable = rownames(arrows_12.SC.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# SC.sigvec_12.wds <- SC.sigvec_12.wds %>% filter(pval < 0.05)
# 
# 
# # Dimensions 1 & 3
# # Fit environmental vectors
# SC.fit_13.wds <- envfit(SC.wds.model, SC.ENV, permutations = 999, choices = c(1,3))
# 
# # Extract vector info
# arrows_13.SC.wds <- as.data.frame(SC.fit_13.wds$vectors$arrows)         # NMDS axis scores
# r2_13.SC.wds     <- SC.fit_13.wds$vectors$r                             # R-squared
# pvals_13.SC.wds  <- SC.fit_13.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# SC.sigvec_13.wds <- arrows_13.SC.wds %>%
#   mutate(R2 = r2_13.SC.wds, pval = pvals_13.SC.wds, Variable = rownames(arrows_13.SC.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# SC.sigvec_13.wds <- SCC.sigvec_13.wds %>% filter(pval < 0.05)
# 
# 
# 
# # Dimensions 2 & 3
# # Fit environmental vectors
# SC.fit_23.wds <- envfit(SC.wds.model, SC.ENV, permutations = 999, choices = c(2,3))
# 
# # Extract vector info
# arrows_23.SC.wds <- as.data.frame(SC.fit_23.wds$vectors$arrows)         # NMDS axis scores
# r2_23.SC.wds     <- SC.fit_23.wds$vectors$r                             # R-squared
# pvals_23.SC.wds  <- SC.fit_23.wds$vectors$pvals                         # p-values
# 
# # Combine into one data frame
# SC.sigvec_23.wds <- arrows_23.SC.wds %>%
#   mutate(R2 = r2_23.SC.wds, pval = pvals_23.SC.wds, Variable = rownames(arrows_23.SC.wds)) %>%
#   select(Variable, everything())
# 
# # Filter for significant vectors (e.g., p < 0.1)
# SC.sigvec_23.wds <- SC.sigvec_23.wds %>% filter(pval < 0.05)
```
 

```{r SC - Model axis scores}
## Extract scores (gives you the coordinates for all three axes)
SC.scores <- scores(LtDC.wds.model, display = "sites")
# head(UDC1.scores)
```

```{r SC - Cat 1 Bio Vars}
## DIMENSIONS 1 and 2
SC.fit1_2.bio <- envfit(SC.wds.model, SC.BIOVEC, permutations = 999, choices = c(1,2))

plot(SC.scores[,1], SC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(SC.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
SC.fit1_3.bio <- envfit(SC.wds.model, SC.BIOVEC, permutations = 999, choices = c(1,3))

plot(SC.scores[,1], SC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
SC.fit2_3.bio <- envfit(SC.wds.model, SC.BIOVEC, permutations = 999, choices = c(2,3))

plot(SC.scores[,2], SC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit2_3.bio, p.max = 0.05)
```
```{r SC - Cat 1 Climate Vars}
## DIMENSIONS 1 and 2
SC.fit1_2.clim <- envfit(SC.wds.model, SC.CLIM, permutations = 999, choices = c(1,2))

plot(SC.scores[,1], SC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(SC.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
SC.fit1_3.clim <- envfit(SC.wds.model, SC.CLIM, permutations = 999, choices = c(1,3))

plot(SC.scores[,1], SC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
SC.fit2_3.clim <- envfit(SC.wds.model, SC.CLIM, permutations = 999, choices = c(2,3))

plot(SC.scores[,2], SC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit2_3.clim, p.max = 0.05)
```

```{r SC - Cat 1 WQ Vars}
## DIMENSIONS 1 and 2
SC.fit1_2.wq <- envfit(SC.wds.model, SC.WQ, permutations = 999, choices = c(1,2))

plot(SC.scores[,1], SC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(SC.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
SC.fit1_3.wq <- envfit(SC.wds.model, SC.WQ, permutations = 999, choices = c(1,3))

plot(SC.scores[,1], SC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
SC.fit2_3.wq <- envfit(SC.wds.model, SC.WQ, permutations = 999, choices = c(2,3))

plot(SC.scores[,2], SC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat1],  
     col = "black", cex = 1.5,
     main="SC: 1-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat1,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit2_3.wq, p.max = 0.05)
```

```{r SC - Cat 3 Bio Vars}
## DIMENSIONS 1 and 2
SC.fit1_2.bio <- envfit(SC.wds.model, SC.BIOVEC, permutations = 999, choices = c(1,2))

plot(SC.scores[,1], SC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 1v2 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(SC.fit1_2.bio, p.max = 0.05)

## DIMENSIONS 1 and 3
SC.fit1_3.bio <- envfit(SC.wds.model, SC.BIOVEC, permutations = 999, choices = c(1,3))

plot(SC.scores[,1], SC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit1_3.bio, p.max = 0.05)

## DIMENSIONS 2 and 3
SC.fit2_3.bio <- envfit(SC.wds.model, SC.BIOVEC, permutations = 999, choices = c(2,3))

plot(SC.scores[,2], SC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 2v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit2_3.bio, p.max = 0.05)
```
```{r SC - Cat 3 Climate Vars}
## DIMENSIONS 1 and 2
SC.fit1_2.clim <- envfit(SC.wds.model, SC.CLIM, permutations = 999, choices = c(1,2))

plot(SC.scores[,1], SC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 1v2 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(SC.fit1_2.clim, p.max = 0.05)

## DIMENSIONS 1 and 3
SC.fit1_3.clim <- envfit(SC.wds.model, SC.CLIM, permutations = 999, choices = c(1,3))

plot(SC.scores[,1], SC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 1v3 dimensions \n Sig bio vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit1_3.clim, p.max = 0.05)

## DIMENSIONS 2 and 3
SC.fit2_3.clim <- envfit(SC.wds.model, SC.CLIM, permutations = 999, choices = c(2,3))

plot(SC.scores[,2], SC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 2v3 dimensions \n Sig Climate vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit2_3.clim, p.max = 0.05)
```

```{r SC - Cat 3 WQ Vars}
## DIMENSIONS 1 and 2
SC.fit1_2.wq <- envfit(SC.wds.model, SC.WQ, permutations = 999, choices = c(1,2))

plot(SC.scores[,1], SC.scores[,2],
     xlab = "NMDS1", ylab = "NMDS2",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 1v2 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(1,2),
            col = color.by.cat,
            draw="polygon", label = T, cex = 0.5);
plot(SC.fit1_2.wq, p.max = 0.05)

## DIMENSIONS 1 and 3
SC.fit1_3.wq <- envfit(SC.wds.model, SC.WQ, permutations = 999, choices = c(1,3))

plot(SC.scores[,1], SC.scores[,3],
     xlab = "NMDS1", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 1v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(1,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit1_3.wq, p.max = 0.05)

## DIMENSIONS 2 and 3
SC.fit2_3.wq <- envfit(SC.wds.model, SC.WQ, permutations = 999, choices = c(2,3))

plot(SC.scores[,2], SC.scores[,3],
     xlab = "NMDS2", ylab = "NMDS3",
     pch = 21, bg = color.by.cat[SC.flow$Cat3],  
     col = "black", cex = 1.5,
     main="SC: 3-month Flow\n 2v3 dimensions \n Sig WQ vectors(p<0.1)");
legend("topright", legend = names(color.by.cat),              
       fill = color.by.cat, title = "Flow", bty = "n");
ordiellipse(SC.wds.model, groups = SC.flow$Cat3,
            choices = c(2,3),
            col = color.by.cat,
            draw = "polygon", label = T, cex = 0.5);
plot(SC.fit2_3.wq, p.max = 0.05)
```
